---
title: "Homework 3"
author: "Emily Duryea"
desription: "Homework 3 submission by Emily Duryea"
date: "12/20/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - homework2
  - emilyduryea
  - student
  - academic
---

# Homework 3

```{r}
# Importing dataset
studentsurvey <- read.csv("_data/student_prediction.csv")
summary(studentsurvey)
```

## Descriptive Statistics & Visualization

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(epiDisplay)

### Sample Information ###

# Sample Gender
tab1(studentsurvey$GENDER, sort.group = "decreasing", cum.percent = TRUE)
# Numeric key: 1 = female, 2 = male

# Sample's Graduated High School Type
tab1(studentsurvey$HS_TYPE, sort.group = "decreasing", cum.percent = TRUE)
# Numeric key: 1 = graduated from a private high school, 2 = state high school, 3 = other)

# Sample's Work Status
tab1(studentsurvey$WORK, sort.group = "decreasing", cum.percent = TRUE)
# Numeric key: 1 = Yes, 2 = No

# Sample's Received Scholarship
tab1(studentsurvey$SCHOLARSHIP, sort.group = "decreasing", cum.percent = TRUE)
# Numeric key: 1 = No Scholarship, 2 = 25% Scholarship, 3 = 50% Scholarship, 4 = 75% Scholarship, 5 = Full Scholarship)

### Research Question 1 ###

# Taking Notes in Class
mean(studentsurvey$NOTES)
median(studentsurvey$NOTES)
sd(studentsurvey$NOTES)
# Numeric key: 1 = never takes notes, 2 = sometimes takes notes, and 3 = always takes notes
# Frequency visualization
tab1(studentsurvey$NOTES, sort.group = "decreasing", cum.percent = TRUE)

# Class Attendance
mean(studentsurvey$ATTEND)
median(studentsurvey$ATTEND)
sd(studentsurvey$ATTEND)
# Numeric key: 1 = always attends class, 2 = sometimes attends class, 3 = never attends class)
# Frequency visualization
tab1(studentsurvey$ATTEND, sort.group = "decreasing", cum.percent = TRUE)

# Reported Listening in Class
mean(studentsurvey$LISTENS)
median(studentsurvey$LISTENS)
sd(studentsurvey$LISTENS)
# Numeric key: 1 = never listens to class lectures, 2 = sometimes listens to class lectures, 3 = always listens to class lectures
# Frequency visualization
tab1(studentsurvey$LISTENS, sort.group = "decreasing", cum.percent = TRUE)

### Research Question 2 ###

# Hours Studying
mean(studentsurvey$STUDY_HRS)
median(studentsurvey$STUDY_HRS)
sd(studentsurvey$STUDY_HRS)
# Numeric key: 1 = 0 hours per week, 2 = <5 hours, 3 = 6-10 hours, 4 = 11-20 hours, 5 = more than 20 hours)
# Frequency visualization
tab1(studentsurvey$STUDY_HRS, sort.group = "decreasing", cum.percent = TRUE)

### Research Question 3 ###

# Peer Study Groups
mean(studentsurvey$PREP_STUDY)
median(studentsurvey$PREP_STUDY)
sd(studentsurvey$PREP_STUDY)
# Numeric key: 1 = studies alone, 2 = studies with friends, 3 = not applicable
# Frequency visualization
tab1(studentsurvey$PREP_STUDY, sort.group = "decreasing", cum.percent = TRUE)

# Positive Class Discussions
mean(studentsurvey$LIKES_DISCUSS)
median(studentsurvey$LIKES_DISCUSS)
sd(studentsurvey$LIKES_DISCUSS)
# Numeric key: 1 = never likes/participates in discussions 2 = sometimes, 3 = always)
# Frequency visualization
tab1(studentsurvey$LIKES_DISCUSS, sort.group = "decreasing", cum.percent = TRUE)
```

### Sample Conclusions

In this particular study, there were more male (60%) than female (40%) participants. Most students attended a state/public high school (71%). Additionally, most students have received at least 50% scholarship at this university (52.4% received 50% scholarship, 29% received 75%, 15.9% received full scholarship, whereas only 2.1% received 25% and 0.7% received no scholarship), indicating that many students at this particular university have received scholarships. Furthermore, most students do not have a job (66.2%) while they are studying at university in this sample. As the vast majority of students have scholarships, working a job during university may not be necessary.

### Research Question Variables Conclusions

For research question 1, most students reported always taking notes (57.3%; m = 2.544828, sd = 0.5649399, median = 3). Most students reported attending class (75.9%; m = 1.241379, sd = 0.429403, median = 1). Interestingly, the majority of students reported only sometimes listening in class (54.2%, m = 2.055172, sd = 0.6747357, median = 2). The statistics reveal that the majority of students engage in classroom engagement behaviors. In regards to research question 2, most students reported studying less than five hours a week (51%; m = 2.2, sd = 0.9174239, median = 2). That is not a lot of time spent studying as I would have anticipated of university students. Most students reported studying alone (73.8%; m = 1.337931, sd = 0.61487, median = 1). Most students also enjoyed class discussions sometimes (48.3%) or all of the time (45.5%). Only a few never enjoyed class discussions and found them beneficial to their learning (6.2%; m = 2.393103, sd = 0.6043425, median = 2).

## Reflection

Limitations of my visuals are that they probably would not be able to be processed by a naive viewer without the numeric key. That is why I provided the necessary numeric values in the code so that it could be understood. Going forward, I would be interested to see how different variables interact with the cumulative GPA (like classroom engagement, peer collaboration, and study habits. Perhaps seeing how demographic variables interact with GPA and other factors (e.g., hours spent studying and work status).
