---
title: "Homework 3 Guanhua Tan"
editor: source
author: "Guanhua Tan"
desription: "Homework 3"
date: "11/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - homework3
  - pisa2018
---

```{r}
# import package
library(tidyverse)
library(ggplot2)
library(tmap)
```

```{r}
# read the data
pisa <- read_csv('_data/CY07_MSU_SCH_QQQ.csv')
dim(pisa)
length(unique(pisa$CNT))
pisa2018 <- pisa %>%
  select(starts_with("SC155"))
dim(pisa2018)
```

# Briefly Describe the Dataset.

This dataset is one part of PISA 2018 dataset with a focus on schools. It covers 80 countries and different regions within each country. The dataset documents 21,903 schools' responses regarding 187 questions. Some key identifiers include CNT (Country Name), STRATUM (Region Name) and OECD (belongs to or not OECD)

# Research Questions

SC115 is a series of questions about the application of digital device in different countries. Some questions measure the availability of digital device in different countries. Some questions interrogate if there are enough supports for teachers in terms of training program, technical support, and time. The dataset categorize the countries into two divisions--OECD or Non-OECD. The dataset reflects regional differences within a country. I'd like to ask a set of questions:What is a general situation all over the world? Is there a difference between OECD countries and Non-Countries? Is there a difference within OECD countries and Non-countries? Are there significant differences regarding different regions within a country? I will present my research results by deploying tables, numbers, and graphics. Furthermore, I will probably explore the reasons that caused the differences.

# Regional Differences all over the World

## Painting the World on a Global Map

```{r}
#tidy data for the world map

pisa2018_joint <- cbind(pisa[,1:12], pisa2018) %>%
  select(CNT, STRATUM, OECD, SC155Q01HA,SC155Q02HA, SC155Q03HA, SC155Q04HA, SC155Q05HA,
         SC155Q06HA, SC155Q07HA, SC155Q08HA, SC155Q09HA, SC155Q10HA, SC155Q11HA)
pisa2018_joint$Accessibility=rowMeans(pisa2018_joint[,c("SC155Q01HA","SC155Q02HA", "SC155Q03HA","SC155Q04HA")])
pisa2018_joint$Human_Resource_Support=rowMeans(pisa2018_joint[,c("SC155Q05HA","SC155Q06HA", "SC155Q07HA","SC155Q08HA","SC155Q09HA", "SC155Q10HA", "SC155Q11HA")])
pisa2018_joint_clean <-pisa2018_joint %>%
  select(CNT, STRATUM, OECD, Accessibility, Human_Resource_Support) %>%
  group_by(CNT) %>%
  mutate(Accessibility_Country_Ave=mean(Accessibility, na.rm=T)) %>%
  mutate(Human_Resource_ave=mean(Human_Resource_Support, na.rm=T)) %>%
  select(CNT,OECD, Accessibility_Country_Ave, Human_Resource_ave) %>%
  distinct() %>%
  arrange(desc(Accessibility_Country_Ave))%>%
  mutate(OECD_or_Non_OECD =case_when(
    OECD==0~"Non-OECD",
    OECD==1~"OECD")) %>%
  select(-OECD)
pisa2018_joint_clean

```

```{r}
# summarize the data
pisa2018_joint_clean %>%
  summary()

```

The dataset covers 80 countries and regions including 21903 schools' responses regrading 187 questions. So I decided to narrow my reach here to five countries. I'd like to look at schools responses regarding a set of questions "SC155". SC155 surveys the accessibility to digital devices and its related training as well as assistance. I clean the data and group_by data as regional code. So I make two different datasets at first. One includes all identifier information. The other covers the columns that contains "SC155Q". Then, I use function "cbind" to combine these two into a new dataset (pisa2018_joint). Furthermore, I found that the questions from SC155Q01HA to SC155Q01HA focus on the accessibility to digital devices while the questions from SC155Q05HA to SC155Q11HA stress on if the schools offer enough training, support, and incentives. I mutate two new variables--"Accessibility" and "Human_Resource_Support" that respectively calculate the average of the former and the latter. Besides, I calculate every country average for "Accessibility" and "Human_Resource_Support" with the function distinct() and rowMeans(). I mutate "OECD_or_Non_OECD" as well in order to create graphic more easily later. After tidying data, the dataset has four columns and 80 rows.

```{r}
# joint data--merge pisa2018 into the world map data
data("World")
world2<-World %>%
  mutate(CNT=iso_a3) %>%
  select(-iso_a3)
world2

world_pisa <-merge(x=world2, y=pisa2018_joint_clean, by="CNT", all.x=T)
world_pisa
```

In order to create a map, I need to use merge function to join the variables I'd like to show in the map dataset "World." I keep all the map dataset information.

```{r}
#draw world map show regional differences
# draw a world map about the accessibly to digital devices and human resource support
tm_shape(world_pisa) +
  tm_polygons(col="Accessibility_Country_Ave",  palette = "Set1")
tm_shape(world_pisa) +
  tm_polygons(col="Human_Resource_ave",  palette = "Set1")
  
```

Finally, I create two world maps reflecting regional differences in terms of the accessibility to digital devices and the human resources support for them all over the world.Apparently, just a few countries have the grades higher than 3 points, which reflects that these countries' agreement of enjoying good accessibility to digital devices.However, even some developed countries still expressed their limited accessibility to digital devices such as United Kingdom and France. The phenomenon, in fact, open the more questions and their answers will require further investigations on these country. I will take France and Unite Kingdom as case studies. With respect to "Human Resource Support," most countries have reported the grades under 3 points. This may not disclose that these countries share the similar scenarios but there is no any objective measurements for them to measure their access to and human resources support on digital devices. Therefore, instead of reflecting that OECD countries lack human resources support, the data may show OECD and non-OECD countries have distinct expectations on human resources support. So self-report can only demonstrate the gap between their expectations and current situations.

## Painting World in Graphics

```{r}
# further tidy data and pivot_longer
# world difference in general

pisa2018_clean_pivot<-pisa2018_joint_clean %>%
  pivot_longer(cols=c(Accessibility_Country_Ave, Human_Resource_ave), 
               names_to = "Group", values_to = "Evaluation")
pisa2018_clean_pivot %>%
   ggplot(aes(Evaluation, fill=Group))+
  stat_boxplot(geom = "errorbar", # Error bars
               width = 0.1)+
  geom_boxplot()+
  facet_wrap(~Group+OECD_or_Non_OECD)+
  labs(title="Pisa2018 Digital Divce Evaluation")+
  coord_flip()

```

In order to further tidy data and create comparative graphics, I first used pivot_longer to create two new variables "Group" and "Evaluation".I put the original variables--"Accessibility_Country_Ave" and "Human_Resource_ave" names to the "Group" and values to "Evaluation." After cleaning and rearranging the dataset, I create four graphics. The boxplot graphics reveal that the OECD countries enjoy the higher access to digital devices than non-OECD countries. But these two groups have reported that their human resource evaluations are in the similar range. The latter reflects the limits to this survey's methodology. Because this survey is reliance on schools' self-report. There is no any objective measurements for them to measure their access to and human resources support on digital devices. Therefore, instead of reflecting that OECD countries lack human resources support, the data may show OECD and non-OECD countries have distinct expectations on human resources support. So self-report can only demonstrate the gap between their expectations and current situations.

## Relation between "Accessibility to Digital Devices" and "Human Resource Support for Digital Human Resource"

```{r}
# Bivariate Visualization--point plot
# suggestion differetiate OECD and Non-OECD
pisa2018_joint_clean %>%
  ggplot(aes(x=Accessibility_Country_Ave, y=Human_Resource_ave, color=OECD_or_Non_OECD))+
  geom_point(size=2)+
  xlab("Accessiblity to Digital Devices for Countires")+
  ylab("Huamn Resourse Support for Countries")+
  geom_smooth(method=lm , color="black", fill="#69b3a2", se=TRUE)
```

The point plot is good at showing the relationship between two groups of numbers. I use it to capture the correlation between "accessibility to digital devices" and"Human Resource Support" in each country. The smooth line has demonstrated the positive correlation between them.

# A further Case Study of Six Countries

I have already deployed a world map and several boxplots to demonstrate regional differences all over the world and arrive at some primitive conclusions. What follows is I'd like to dig into 6 countries ( Hong Kong, Philippines, Argentina and Brazil belong to non-OECD, Britain and France belong to non-OECD).

```{r}
# tidy data about the case study
# choose five countries to look at regional difference within the countries
# United Kingdom (GBR), Hong Kong (HKG), Philippines (PHL), Argentina (ARG), Brazil(BRA), France(FRA)
pisa2018_joint_case_study <- pisa2018_joint  %>%
  select(CNT, STRATUM, OECD, SC155Q01HA,SC155Q02HA, SC155Q03HA, SC155Q04HA, SC155Q05HA,
         SC155Q06HA, SC155Q07HA, SC155Q08HA, SC155Q09HA, SC155Q10HA, SC155Q11HA) %>%
  group_by(STRATUM) %>%
  arrange(STRATUM) %>%
  filter(CNT== "GBR" | CNT== "HKG" | CNT=="PHL"| CNT=="ARG" | CNT=="BRA" |CNT=="FRA" )
pisa2018_joint_case_study
```

```{r}
#further tidy data

pisa2018_joint_case_study$Accessibility=rowMeans(pisa2018_joint_case_study[,c("SC155Q01HA","SC155Q02HA", "SC155Q03HA","SC155Q04HA")])
pisa2018_joint_case_study$Human_Resource_Support=rowMeans(pisa2018_joint_case_study[,c("SC155Q05HA","SC155Q06HA", "SC155Q07HA","SC155Q08HA","SC155Q09HA", "SC155Q10HA", "SC155Q11HA")])
pisa2018_case_study_clean <- pisa2018_joint_case_study %>%
  select(CNT,STRATUM,OECD, Accessibility, Human_Resource_Support) %>% 
  pivot_longer(cols=c(Accessibility, Human_Resource_Support), names_to = "Group", values_to = "Evaluation")%>%
   mutate(OECD_or_Non_OECD =case_when(
    OECD==0~"Non-OECD",
    OECD==1~"OECD")) %>%
  select(-OECD)
pisa2018_case_study_clean
```

Following the above-mentioned finding that the questions from SC155Q01HA to SC155Q01HA focus on the accessibility to digital devices while the questions from SC155Q05HA to SC155Q11HA stress on if the schools offer enough training, support, and incentives. I apply the same tidy data approach to the case study's dataset. I mutate two new variables--"Accessibility" and "Human_Resource_Support" that respectively calculate the average of the former and the latter. I just calculate the averages of these two groups of every school (4228 schools). I use pivot_longer to rearrange the dataset in order to show these two categories in one graphic. I The result is the dataset "pisa2018_case_study_clean."

```{r}
# violin graphic
pisa2018_case_study_clean %>%
  ggplot(aes(CNT,Evaluation, fill=CNT), na.rm=T) +
  geom_violin()+
  facet_wrap(~Group)+
  xlab("Country")+
  ylab("Evaluation")+
  ggtitle("Schools' Evaluation on Digtial Device") 

```

I use the violin graphics to disclose the evaluation of digital devices in six countries with the dimension of "Accessibility" and "Human Resource Support". The graphic has shows, whatever accessibility and human resource support, Hong Kong, France and Britain report higher grades than Argentina, Brazil, and the Philippines. most schools' in Hong Kong, France and Britain response are around the range of "agree" while most schools' in Argentina, Brazil, the Philippines replies are around the range of "disagree", which reflects that schools in Argentina, Brazil, the Philippines are not satisfied by their current situation.

```{r}
pisa2018_case_study_clean %>%
  ggplot(aes(CNT, Evaluation,fill=CNT), na.rm=T)+
  geom_violin()+
  xlab("Country")+
  ylab("Evaluaiton")+
  ggtitle("Schools' Evaluation on Digtial Device")+
  facet_wrap(~Group+OECD_or_Non_OECD)
```

I add one more dimension--OECD or Non-OECD to this violin plot. In terms of "Accessibility," Hong Kong has a better access to digital devices than Argentina, Brazil, the Philippines. the similar situation happens to the evaluation of "Human Resources Support." The fact may be interpreted that Hong Kong has more enough fiscal resources to support their schools. In fact, the plot demonstrates the variation of the evaluation on digital devices in non-OECD countries. By contrast, two OECD countries--France and Brain share the similar shape in terms of accessibility and human resource support, which reflects that most schools in these two countries believe that they enjoy an good access to digital device and receive enough human resources support.

```{r}
pisa2018_joint <- cbind(pisa[,1:12], pisa2018) %>%
  select(CNT, STRATUM, OECD, SC155Q01HA,SC155Q02HA, SC155Q03HA, SC155Q04HA, SC155Q05HA,
         SC155Q06HA, SC155Q07HA, SC155Q08HA, SC155Q09HA, SC155Q10HA, SC155Q11HA) 
```

```{r}
df<-pisa2018_joint %>%
  group_by(CNT) %>%
  filter(CNT== "GBR" | CNT== "HKG" | CNT=="PHL"| CNT=="ARG" | CNT=="BRA" |CNT=="FRA" ) %>%
  drop_na() %>%
  count(Q1=factor(SC155Q01HA))%>%
  mutate(pct=paste0(sprintf("%4.1f",n/sum(n)*100), "%"))

ggplot(df, aes(x=CNT, y=n ,fill=Q1), na.rm=T)+
  geom_bar(stat="identity", position="dodge")+
  geom_text(aes(label=pct), 
            position = position_dodge(width = 1),
            vjust=-0.5, size=2)+
  ylab("Schools' Response Count")+
  xlab("Country")+
  labs(fill="Digital Device Evaluation")+
  scale_fill_discrete(labels=c("Strongly Disagree", "Disagree", "Agree", "Strongly Agree"))
```

Similar to the violin graphics, the bar plot shows the homogeneous tendency. more responses from France, Britain and Hong Kong reported "disagree" that that from Argentina, Brazil, and the Philippines. At the same time, the bar plot discloses the limits of the dataset pisa2018. The number of responses from France is relatively fewer in terms of its size. In other words, the sample of France is very tiny, which may not reflect the situation there, compared with the samples from the rest five countries and area.

# A Brief Conclusion and Reflection

The world map and box plots have demonstrated that concerning the accessibility to digital devices, most of OECD countries are satisfied by their accessibility. By contrast, most of Non-OECD countries are disappointed by it. I am surprised by the finding that both these groups complained their "human resources support." However, this fact reflects the limits of the survey's methodology that there is no objective measurement for these countries to measure their human resources support. The case studies indicates the similar distribution of schools' response within Argentina, Brazil, the Philippines. Hong Kong, France and UK share the another similar distribution. Although it is not a member of OECD, Hong Kong share more characteristics with OECD countries. I am wondering how can I offer clearer descriptions and attractive narrative for graphics and how can I find better graphics to present the data. Furthermore, I have noted that the dataset and my visualization only present the differences of eighty countries' accessibility to and human resources support for digital devices but fail to explain what cause the differences. Besides, even two countries present the same number. It may not show these two countries share the similar circumstance. As I mentioned before, numbers here only reflect the gap between current experiences and expectations of every country. This is the limits on our survey approach.
