---
title: "Challenge 1 Solution Alex Gauvin-Valenta"
author: "Alex Gauvin-Valenta"
desription: "Reading in data and creating a post"
date: "9/26/22"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
railroad<-read_csv("_data/railroad_2012_clean_county.csv")


```

## Summary

The railroads data set looks to be showing the total number of railroad employees per county in each state. Cook, IL has the most employees at 8,207 while Sitka, AK has the least at 1.


```{r}
#| label: summary
summary(railroad)

#Data Table
railroad%>%
  arrange(total_employees,desc(total_employees))
```
