---
title: "Jayson Tatum: A Future Hall of Famer?"
description: "predicting if Jayson tatum will have a Hall of Fame Career"
author: "Amer Abuhasan"
date: "`r Sys.Date()`"
output: distill::distill_article
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(plyr)
library(tidyverse)
library(rvest)
library(ggrepel)
library(readr)
library(RCurl)
library(jpeg)
library(readxl)
library(dplyr)
library(scales)
library(ggplot2)
library(knitr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, paged.print=TRUE)
```

# Introduction

The NBA is one of the most popular leagues in the world.It is filled
with a massive amount of excitement, heart breaks and history. From
coast to coast, fans across the world tune in game after game to watch
their favorite teams and players perform. The sport draws many debates
between fan bases and fan favorites. In the NBA there are hundreds of
cemented legendary players to step foot on the parcel and dedicate their
lives to the sport they love. Some of the Best players to be named are
Lebron James and Wilt Chamberlain, who both are considered one of the
best offensive players in the sports history. Today We see lots of young
new superstars who are expected to be generational players.

How do we compare these new generational players to those of the past?
The NBA is a point based league, the more points you put up the better
chances the player has to be ranked amongst the greats. For this
research I decided to use one young generational player in Jayson
Tatum(Boston Celtics), and two from two different past Generations in
Lebron James(LA Lakers) and Wilt Chamberlain(retired). Lebron is
considered the best of all time, while Chamberlain is considered one of
the best offensive scoring players of all time. Since Jayson Tatum is in
his fifth season, it is only fitting that I compare the first 5 seasons
between all the players' offensive statistics.

# Research Question

The goal is to comprehend Jayson's Offensive stats and compare them to
Lebron James and Chamberlain's in order to see if Jayson will have a
Hall of Fame career. Since Lebron has not retired yet, I would like to
compare him to Jayson Tatum since he is directly the generation before
Tatum's. I will then Compare Jayson to Wilt which is a few generations
in the past. Also choosing Wilt is a great comparison because Wilt
scored the most points in the first 5 seasons than any other player.

Can we find a trend that indicates that Jayson Tatum will be in the
conversation with the greatest players of all time?

# Data

The website where the data was collected from (Basketball Statistics &
History). Basketball reference was created in 2004, the company teamed
up with stat reference.com and has thousands of statistics on players in
most sports. The data used is accurate and is easy to download into csv
files in order to import the data into R.

# **Data Reading**

```{r message=FALSE, paged.print=TRUE}
library(readxl)
Jayson_Tatum_career_stats <- read_excel("_data/Jayson_Tatum_career_stats.xlsx")
kable(
 
Jayson_Tatum_career_stats)

#insert lebron James stats/skip line for correct columns 
Lebron_james_stats <- read_excel("_data/Lebron_james_stats.xlsx",skip = 1)

kable(
Lebron_james_stats)

```

The table above is the original data for Jayson
Tatum(Jayson_Tatum_career_stats) and lebron
james(lebron_James_career_stats). The raw data has a lot of unnecessary
statistics(Defensive).

# **Trimming and data cleaning** 

Since we are only looking at offensive stats, the only option was to
trim the data and tidy it to columns we only need.

The select(c() allowed for the ability to select the column names we
needed, this offered a cleaner look at the offensive data for
Jayson,James and Chamberlain. Renaming the Data frame to
X_relevant_stats

I created two different sub data sets for each of the 2 players to
separate columns with percentages and columns with decimals as well. The
two sub datasets for each player contain main statistics and the other
include shooting statistics. The tables were labeled (x_Shooting_stats),
(x_Main_stats).

```{r message=FALSE, paged.print=TRUE}
# make a list of stats we care about
Relevant_stats=Jayson_Tatum_career_stats %>%
  select(c("Season","Age","Tm","G","MP","FG","FGA", "PTS","2P","2P%","3P","3P%","FT","FT%","FG%","AST"))
(kable(Relevant_stats))


#Create 2 Sub Data sets by selecting stats type for Jayson Tatum 
Shooting_stats<- Relevant_stats %>% select(c("Season","Tm","2P%","3P%","FT%","FG%"))
(kable(Shooting_stats))

Main_stats<-Relevant_stats %>% select(c("Season","Tm","MP","FG","FGA", "PTS","AST","G"))
(kable(Main_stats))
  
# make a list of stats we care about for lebron james 
Lebron_Relevant_stats=Lebron_james_stats %>%
  select(c("Season","Age","Tm","MP", "G","FG","FGA","PTS","2P","2P%","3P","3P%","FT","FT%","FG%","AST"))



(kable(Lebron_Relevant_stats))
#Create 2 Sub Data sets by selecting Stats type for Lebron James 
Lebron_Shooting_stats<- Lebron_Relevant_stats %>% select(c("Season","Tm","2P%","3P%","FT%","FG%"))
(kable(Lebron_Shooting_stats))

Lebron_Main_stats<-Lebron_Relevant_stats %>% select(c("Season","Tm","MP","FG","FGA", "PTS","AST", "G"))

(kable(Lebron_Main_stats))
```

```{r message=FALSE, paged.print=TRUE}
#allow for decimals to convert to percentages
myfun <- function(x) {
  if(is.numeric(x)){ 
    ifelse(is.na(x), x, paste0(round(x*100L, 1))) 
  } else x 
}
```

The Column name "Year" was not included in the original data set, I
included a code to rank by year because the players did not all start
their NBA careers at the same time, the Column "Season" was not feasible
to rank the players alongside each other. Thus, the reasoning for the
code below, I mutate the data to add a new column "Year" with ranking
from 1-5

```{r message=FALSE, paged.print=TRUE}

#make stats percentages 
Shooting_stats<-Shooting_stats%>% mutate_each(funs(myfun))

Lebron_Shooting_stats<-Lebron_Shooting_stats%>% mutate_each(funs(myfun))


#Selecting first 5 years for lebron and Jayson Tatum 
Shooting_stats<-head(Shooting_stats,5)%>%mutate(player_name="Jayson Tatum",Year=c(1:5))

Main_stats<-head(Main_stats,5)%>%mutate(player_name="Jayson Tatum",Year=c(1:5))

Lebron_Shooting_stats<-head(Lebron_Shooting_stats,5)%>% mutate(player_name="Lebron James",Year=c(1:5))

Lebron_Main_stats<-head(Lebron_Main_stats,5)%>% mutate(player_name="Lebron James",Year=c(1:5))
```

# Connecting the data between Lebron and Jayson

In order to compare the Two players and set them up for charts and
graphing, I needed to merge the different data sets that were cleaned
for the players. Instead of the merge function I used rbind(). rbind()
can be used to bind or combine several vectors, or data frames by rows.

It is easier to look at side by side and includes player names just
incase the player switches teams, which in this case did not happen but
will happen when we add Wilt Chamberlain

```{r message=FALSE, paged.print=TRUE}

#merge by using rbind to compare the different stats between the 2 players 
Shooting_stats_comparision<-rbind(Shooting_stats,Lebron_Shooting_stats )

Main_stats_comparision<- rbind(Main_stats,Lebron_Main_stats)
```

# **visualization**

\<-Using the first code below allowed for the creation of a gg plot that
shows FG% between years of Jayson and Lebron. The graph shows a clear
visualization comparing Lebrons and Tatum's Shooting percentages. We can
see that in Tatum's first year he had a very high shooting % meanwhile
Lebron not so much. Lebron Continued to trend upwards while Tatum
Trended Down.

\<- The second line of code created another GG line plot but this time
For PTS by year. If you pay attention, You can realize that something
doesn't add up between FG% and points, Especially comparing each other's
first season. Tatum had a higher FG% than Lebron and still had way fewer
made PTS. I wanted to see what other variables may have contributed to
this cause.

```{r message=FALSE, paged.print=TRUE}
#creeating a GG plot(line chart) for shooting stats using FG% between the two players 
Shooting_stats_comparision %>%
  ggplot( aes(x=Year, y=Shooting_stats_comparision$`FG%`, group=player_name,color=player_name)) +
    geom_line() +
    geom_point()

#Creating GG plot(line chart) for main stats using PTS between the two players 
Main_stats_comparision %>%
  ggplot( aes(x=Year, y=PTS, group=player_name,color=player_name)) +
    geom_line() +
    geom_point()
```

# Creating theories 

I decided to see playing time(MP) and decipher if this has a substantial
effect on why Tatum is not scoring as much, despite the high FG%. The
code below allows for a creation of a box plot.Box Plots are a great
chart to use when showing the distribution of data points across a
selected measure.

The First plot indicates that Lebron James playing time is much Higher
than Jayson Tatums, We can also see the median between the two for
minutes played through the first 5 years.

The second plot(Bar) builds a case clarifying that Playing more minutes
actually does result in more points especially for Jayson Tatum since
his playing minutes went up tremendously since his first year. Seeing
this chart most definitely indicates that Lebron is scoring more points
due to minutes, but I want to see what other variables allow for PTS to
be higher due to MP. We can see in Year 4 that Jayson with less minutes
played(not by much) had almost equivalent points.

[**What comes from playing more minutes?**]{.underline}

```{r message=FALSE, paged.print=TRUE}

Main_stats_comparision%>%
  ggplot( aes(x=factor(player_name), y=MP, fill=player_name)) +
    geom_boxplot() +
    ggtitle("MP comparision") +
    xlab("")


Main_stats_comparision %>%
  ggplot( aes(x=MP, y=PTS,fill=player_name)) + 
   geom_bar(stat = "identity") +
  facet_wrap(~Year)


```

Since Tatums points went up with more playing time, we can make an
assumption that something changed, lets test out some variables that
contribute to points(PTS) and see if there is any corolation.

# **Analyzing different variables**

The two codes will graph us

\<-A: FGA(field goal attempts) by MP

\<-B : FGA with FG (field goal made). I attempted this to study the
charts and analyze what the graphs are trying to say.

```{r message=FALSE, paged.print=TRUE}

#A
Main_stats_comparision %>%
  ggplot( aes(x=FGA, y=MP, group=player_name,color=player_name)) +
    geom_line() +
    geom_point()
#B
Main_stats_comparision %>%
  ggplot( aes(x=FGA, y=FG, group=player_name,color=player_name)) +
    geom_line() +
    geom_point()

```

What are we seeing here?

In graph A it is showing that the more MP the more FGA Jayson is taking.
For lebron it is fair to say that his minutes are already so
high(39-42.5MP) so his FGA are always high(about 18-24) shots per game.
Hence there is only 60 minutes in a game.

In graph B We can see that with more FGA the more FG that are made for
Jayson Tatum, the same goes for Lebron, So I will conclude that Tatum
needs more minutes in order to take more FGA and that raises the chance
of FG made.

# Adding third comparision

Inserting Wilt to see if theory is not a fluke and to see trends. Now
for the steps we did early on, we need to (1)Read excel (2)select(c)
data we need from massive table (3)creating the two subdata shooting and
main stats tables (4)mutating and creating column "Year" so we can
compare them since they are in different generations. (5)Rbind all the
sub data into one main data between all three players (6)create graphs
and analyze trends with the third player.

```{r message=FALSE, paged.print=TRUE}


library(readxl)
Wilt_chamb_stats <- read_excel("_data/Wilt_chamb_stats.xlsx")
Wilt_chamb_stats
kable(Wilt_chamb_stats)


Wilt_Relevant_stats= Wilt_chamb_stats %>% 
  select(c("Season","Age","Tm","G","MP","FG","FGA", "PTS","FT","FT%","FG%","AST"))


kable(Wilt_Relevant_stats)

#select stats type for Wilt
Wilt_Shooting_stats<- Wilt_Relevant_stats %>% select(c("Season","Tm","FT%","FG%"))
kable(Wilt_Shooting_stats)

Wilt_Main_stats<-Wilt_Relevant_stats %>% select(c("Season","Tm","MP","FG","FGA", "PTS","AST","G"))
kable(Wilt_Main_stats)
#Making wilts stats into percentages 
Wilt_Shooting_stats<-Wilt_Shooting_stats%>% mutate_each(funs(myfun))

#Comparing them by year so number 1-5
Wilt_Shooting_stats<-head(Wilt_Shooting_stats,5)%>%mutate(player_name="Wilt Chamberlain",Year=c(1:5))

Wilt_Main_stats<-head(Wilt_Main_stats,5)%>%mutate(player_name="Wilt Chamberlain",Year=c(1:5))
kable(Wilt_Main_stats)
      
kable(Wilt_Shooting_stats)



Ultimate_Shooting_stats<-rbind.fill(Shooting_stats,Lebron_Shooting_stats,Wilt_Shooting_stats )

Ultimate_Main_stats<- rbind.fill(Main_stats,Lebron_Main_stats,Wilt_Main_stats)
kable(Ultimate_Main_stats)
kable(Ultimate_Shooting_stats)

```

# Analyze trend

Analyzing the graphs: we can see that that when including Wilt, since he
is the #1 most scoring player in the first 5 seasons of any NBA
career(statistical Fact), Wilts MP is super high and So is his points,
Wilt was almost putting up a point per minute! Unreal stat. You can
notice a trend with all 3 players that the more Minutes they play the
more Points they score. Jayson Tatum as we talked about earlier is
gaining more minutes per season which is causing him to score more
points, Ultimatley pushing him the right direction.

```{r message=FALSE, paged.print=TRUE}
#comparing the 3 players PTS by Minutes played 
Ultimate_Main_stats %>%
  ggplot( aes(x=MP, y=PTS,fill=player_name)) + 
   geom_bar(stat = "identity") +
  facet_wrap(~Year)

#comparing the 3 players PTS By Field goal attempts
Ultimate_Main_stats %>%
  ggplot( aes(x=FGA, y=PTS,fill=player_name)) + 
   geom_bar(stat = "identity") +
  facet_wrap(~Year)

#comparing the 3 players FG made by Field Goal Attemts 
Ultimate_Main_stats %>%
  ggplot( aes(x=FGA, y=FG,fill=player_name)) + 
   geom_bar(stat = "identity") +
  facet_wrap(~Year)

```

# Conclusion

For Tatum to be considered one of the greats, he needs to get more
playing time, which his trajectory looks like he is trending in the
correct direction. We now know that with more minutes played, comes more
opportunities to shoot the ball and the more you shoot the ball the more
you will make, obviously Tatums FG% is high as we noted earlier, He is
looking like he is going to be a star in this league.

[**Keeping up to date on Jaysons 6th and current ongoing
season.**]{.underline}

Tatum is in the middle of completing his 6th season, through 27 games so
far he is averaging 36.7 MP, 9.8 FG made,21 FGA, FG% 0.468%, 29.7 PTS.
All those stats are higher than the previous seasons, His team is
currently first in the NBA and he is the front runner for the MVP(most
valuable player) award thus far, according to (Kia MVP Ladder).

# bibliography

1-Basketball Statistics & History of every Team & NBA and WNBA players.
Basketball. (n.d.). Retrieved December 14, 2022, from
https://www.basketball-reference.com/

2-Jayson Tatum stats, height, weight, position, draft status and more
(no date) Basketball. Available at:
https://www.basketball-reference.com/players/t/tatumja01.html.

3-"Lebron James Stats, Height, Weight, Position, Draft Status and More."
Basketball,
https://www.basketball-reference.com/players/j/jamesle01.html.

4- Wilt Chamberlain Stats, Height, Weight, Position, Draft Status and
more. (n.d.). Basketball-Reference.com.
https://www.basketball-reference.com/players/c/chambwi01.html

5- R Core Team (2022). R: A language and environment for statistical
computing. R Foundation for Statistical Computing, Vienna, Austria. URL
https://www.R-project.org/.

6- Wright, Michael C. \"Kia MVP Ladder: Giannis and Luka battle, Morant
surges into Top 5.\" NBA.com, 9 December 2022,
https://www.nba.com/news/kia-mvp-ladder-dec-9-2022-edition. Accessed 17
December 2022.
