---
title: "Challenge 2 Kristin Abijaoude"
author: "Kristin Abijaoude"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - kristin_abijaoude
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
railroads <- read_csv("_data/railroad_2012_clean_county.csv")
railroads
```

This dataset is county-level data from censuses across 2930 counties with at least one railroad employee. Each row has information about the county name, the state that the county is in, and the total number of railroad employees in said county. The information is likely collected from the 2012 US Census.
```{r}
#| label: Summary
summary(railroads)
```

```{r}
#| label: Select ()
## Selecting county and total employees
select(railroads, `county`, `total_employees`)
```
```{r}
#| label: Mean ()
## Average # of employees in each state
railroads %>%
  group_by(state) %>%
  summarise(mean.employees = mean(total_employees, na.rm = TRUE))
```

```{r}
#| label: filter ()
## filtering counties with up to 100 railroad employees
filter(railroads,`total_employees` <= 100)
```

```{r}
#| label: the big stuff
summarize(railroads, mean.total_employees = mean(`total_employees`, na.rm = TRUE), median.total_employees = median(`total_employees`, na.rm = TRUE), min.total_employees = min(`total_employees`, na.rm = TRUE), max.total_employees = max(`total_employees`, na.rm = TRUE), sd.total_employees = sd(`total_employees`, na.rm = TRUE), var.total_employees = var(`total_employees`, na.rm = TRUE), IQR.total_employees = IQR(`total_employees`, na.rm = TRUE))
```
There are many more counties with a lower number of total railroad employees than those with a higher number of them, which partially explains the low numbers for median and mean, and why the mean is greater than the median. The standard deviation is very high which shows there were outliers with large values, such as Cook County, IL with 8207 railroad employees. 

```{r}
#| label: Visualizing Population Density
library(ggplot2)
ggplot(railroads, aes(x=state, y=total_employees)) + 
    geom_point()
```
I attempted to create a scatterplot graph to display the amount of railroad employees by state, with each dot representing county. Plenty of states have counties that are more clustered, while some states have one or outlier counties, such as Illinois and New York. When you look into population density of the those states, you can see that the majority of the population lives in a particular area, while the rest of the state is relatively sparse.
