---
title: "Challenge 2 Instructions"
author: "Meredith Rolfe"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - faostat
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xls ⭐
-   FAOstat\*.csv or birds.csv ⭐⭐⭐
-   hotel_bookings.csv ⭐⭐⭐⭐

```{r}
# download library

library(tidyverse)

# read in data
# note that R does not recognize '\' --> needs to be '/' or '\\'

birds <- ("C:\\Users\\caitr\\OneDrive\\Documents\\DACSS\\601_Fall_2022\\posts\\_data\\birds.csv")
read_csv(birds)

birds<-read_csv("_data\\birds.csv")

getwd()

birds

# try reading in using just file name
# keep receiving error messages (no such file or directory), (object not found), (cannot open the connection)
# will continue to read in using entire pathway
```

Add any comments or documentation as needed. More challenging data may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: summary
# This data set contains 30,977 rows and 14 columns.
# The data includes information about different types of birds,
# including their relative geographic location and their species.
# Want to gather more information about column names and row-level data.

# show column names

colnames("C:\\Users\\caitr\\OneDrive\\Documents\\DACSS\\601_Fall_2022\\posts\\_data\\birds.csv")
colnames(birds)

# NULL 
# try 'names' to get character vector

names("C:\\Users\\caitr\\OneDrive\\Documents\\DACSS\\601_Fall_2022\\posts\\_data\\birds.csv")

# NULL
# identify missing values

is.na("C:\\Users\\caitr\\OneDrive\\Documents\\DACSS\\601_Fall_2022\\posts\\_data\\birds.csv")

# FALSE
# try summary

summary(birds)

# try colnames with 'test'

colnames('test')

# NULL

colnames(birds)
names(birds)

# NULL

data.frame(birds)
colnames(birds)
names(birds)
colnames("C:\\Users\\caitr\\OneDrive\\Documents\\DACSS\\601_Fall_2022\\posts\\_data\\birds.csv")

# NULL

# use 'spec()' to get full column specification

spec(birds)

# The columns in this data set are named: 'Domain Code,' 'Domain,' 'Area Code,' 'Area,' 'Element Code,' 'Element,' 'Item Code,' 'Item,' 'Year,' 'Unit,' 'Value,' 'Flag,' and 'Flag Description.'

# re-ran 'spec(birds)', now it's not working
# does 'nrow' or 'ncol' work?

ncol(birds)

# NULL

nrow(birds)

# NULL

# needs to be a tibble created through 'readr'

library(readr)
class(birds)
readr::spec(as_tibble(birds))

# NULL

spec(birds)

# same error message: Error in spec(birds) : inherits(x, "tbl_df") 
# is not TRUE

```

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

```{r}
# remove columns containing 'code'

library(dplyr)%>%
select(birds, contains("Code"))

# no applicable method for 'select' applied to an object of class "character"
# need to create data frame or tibble

# create a tibble

tbl_df(birds)
as_tibble(birds)

# tibble is displaying as 1x1

read_csv(birds)
birds%>%select(contains("Code"))
select(birds, -ends_with("Code"))

# keep needing to reinstall library and packages?

install.packages("dplyr")
library(tidyverse)

read_csv("C:\\Users\\caitr\\OneDrive\\Documents\\DACSS\\601_Fall_2022\\posts\\_data\\birds.csv")
birds<-"C:\\Users\\caitr\\OneDrive\\Documents\\DACSS\\601_Fall_2022\\posts\\_data\\birds.csv"

select(birds, -ends_with("Code"))
dplyr::select(-ends_with("Code"))

as_tibble(birds)
dplyr::select(-ends_with("Code"))
dplyr::select(contains("Code"))

# no matter what I do, I keep getting the 'no applicable method for 'select' applied to an object of class "character"'
```

### Explain and Interpret

Be sure to explain why you choose a specific group. Comment on the interpretation of any interesting differences between groups that you uncover. This section can be integrated with the exploratory data analysis, just be sure it is included.
