---
title: "Challenge 6 Instructions"
author: "Meredith Rolfe"
description: "Visualizing Time and Relationships"
date: "08/23/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_6
  - hotel_bookings
  - air_bnb
  - fed_rate
  - debt
  - usa_households
  - abc_poll
  
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  create at least one graph including time (evolution)
   - try to make them "publication" ready (optional)
   - Explain why you choose the specific graph type
5)  Create at least one graph depicting part-whole or flow relationships
   - try to make them "publication" ready (optional)
   - Explain why you choose the specific graph type

[R Graph Gallery](https://r-graph-gallery.com/) is a good starting point for thinking about what information is conveyed in standard graph types, and includes example R code.

(be sure to only include the category tags for the data you use!)

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

  - debt  ⭐
  - fed_rate ⭐⭐
  - abc_poll ⭐⭐⭐
  - usa_hh ⭐⭐⭐
  - hotel_bookings ⭐⭐⭐⭐
  - AB_NYC ⭐⭐⭐⭐⭐

  

```{r}

ab_nyc <- read_csv("_data/AB_NYC_2019.csv")
ab_nyc
summary(ab_nyc)

```

### Briefly describe the data

I chose to work with the NYC AirBNB data. It consists of 16 variables for 48,895 different AirBNB listings. The variables consist of listing name and ID, host name and ID, neighborhood and neighborhood_group(borough), latitude and longitude, room type, price, min nights, number of reviews, last review, reviews per month, host listings count, and days of availability.

## Tidy Data (as needed)

I took a look at the summary statistics for the data on the read in. A quick observation I made was that there were 10052 AirBNBs with no review. This is something that I want to filter out so I can accurately make a graph based on review date.

```{r}

new_abnyc <- filter(ab_nyc, `number_of_reviews` > 0)
new_abnyc
summary(new_abnyc)

```

Looking at the newly filtered data, I still see some things that give me questions. The max reviews per month is 58.5, or almost two reviews per day, which seems impossible. There is also a minimum number of nights of 1250 for a listing, also very questionable. Over a quarter of the listings have 0 days of availability and I am not sure if that will affect anything or not.

```{r}

rev_bym <- new_abnyc%>%
  arrange(desc(`reviews_per_month`))
rev_bym

```

It appears to be an outlier, but there is nothing out of the ordinary so I will leave it alone. 

```{r}

nights <- new_abnyc%>%
  arrange(desc(`minimum_nights`))
nights

no_av <- filter(new_abnyc, `availability_365` == 0)
no_av
count(no_av, `room_type` == "Entire home/apt")
```

There are a few listings with a high amount of minimum nights. Most are entire homes/apartments, so I am guessing that these are just long term living spaces. I also wanted to see if the all the listings with 0 availability for the year meant anything significant. I found out that half of the listings with no availability were Entire home/apt, which is also consistent with a long term living space.


##Are there any variables that require mutation to be usable in your analysis stream? For example, do you need to calculate new values in order to graph them? Can string values be represented numerically? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?


```{r}

```

## Time Dependent Visualization

My first visualization is to represent the most current review dates in a density chart. I felt like this is the best visual way to represent the reviews over time combined with the quatity of data.

```{r}
review_dates <- new_abnyc %>%
  ggplot(aes(x = `last_review`)) +
  geom_density(fill="cyan") +
  labs(x="Most Recent Review Date", title = "Density of Latest Review Dates for NYC AirBNB") +
  theme_dark()
review_dates
```
To get a better visual for the data I filtered the data so that it only shows from 2015 onward.

```{r}

review_dates_2015 <- new_abnyc %>%
  filter(`last_review` >= "2015-01-01")%>%
  ggplot(aes(x = `last_review`)) +
  geom_density(fill="cyan") +
  labs(x="Most Recent Review Date", title = "Density of Latest Review Dates for NYC AirBNB") +
  theme_dark()
review_dates_2015

```

## Visualizing Part-Whole Relationships

```{r}

```
