---
title: "Challenge 6"
author: "Lai Wei"
description: "Visualizing Time and Relationships"
date: "11/19/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_6
  - hotel_bookings
  - air_bnb
  - fed_rate
  - debt
  - usa_households
  - abc_poll
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  create at least one graph including time (evolution)
   - try to make them "publication" ready (optional)
   - Explain why you choose the specific graph type
5)  Create at least one graph depicting part-whole or flow relationships
   - try to make them "publication" ready (optional)
   - Explain why you choose the specific graph type

[R Graph Gallery](https://r-graph-gallery.com/) is a good starting point for thinking about what information is conveyed in standard graph types, and includes example R code.

(be sure to only include the category tags for the data you use!)

## Read in data

  - abc_poll ⭐⭐⭐

```{r}
rate <- read_csv("_data/abc_poll_2021.csv")
rate
```

### Briefly describe the data

## Tidy Data (as needed)

Get the colnums name. 
```{r}
colnames(rate)
```
From this section, I made a new proportion table to show the demographic information, races percentage in this survey. Obviously, the White, non-Hispanic is the largest group.  
```{r}
rate_1 <- rate %>% 
  select(xspanish, starts_with("pp")) 
prop.table(table(rate_1$ppethm))
```



```{r}
rate_ethm <- rate %>% 
  mutate(Ethm = ifelse(ppethm == "White, Non-Hispanic", "non-ethnic minorities", "ethnic minorities"))

rate_ethm
```

## Time Dependent Visualization

In this section, I made a graphs by ggplot in different ages group. And we can see even though the age range is wide, the data distribution is quite similar. People's opinion do not change too much as time going by. 
```{r}
rate %>% 
  ggplot(aes(Q2)) + geom_bar() + theme_bw() + 
  facet_wrap(vars(ABCAGE), scales = "free")
```

## Visualizing Part-Whole Relationships

After adding fill section in ggplot, it shows different colors on every bar based on respondents' political tendency. 
```{r}
rate %>% 
  ggplot(aes(Q2, fill = QPID)) + geom_bar() + theme_bw() + 
  facet_wrap(vars(ABCAGE), scales = "free")
```

