---
title: "Challenge 1"
author: "Tracy Tien"
desription: "Reading in data and creating a post"
date: "What is time these days?"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
birds <- read_csv("_data/birds.csv")
spec(birds)
# birds.csv has 14 columns, and the results of the column specifications indicates that the first line of the .csv contains column names.
dim(birds)
# birds.csv has 30977 rows and 14 columns
head(birds)
colnames(birds)
```


Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
# Some of the columns are intuitive (i.e. Area = country, Unit = quantity) but others can be explored further.
table(select(birds, "Domain Code")) 
table(select(birds, "Domain"))
# There's only 1 "Domain" type: Live Animals, and the code is: QA
table(select(birds, "Area"))
n_distinct(select(birds, "Area"))
# Based on the "Area" and "Flag Description," which there are 248 UN Food and Agriculture recognized countries 
table(select(birds, "Item"))
# The types of birds under items are: chickens, Ducks Geese and guinea fowls, pigeons/other birds, turkeys
table(select(birds, "Year"))
# The records span from 1961 to 2018.
# To summarize, there are 248 countries (under UN FAO), from 1961-2018, where "Value" indicates each year's total value for each type of livestock bird (chicken, ducks geese and guinea fowls, pigeons and other birds, turkeys)
```
