---
title: "Challenge 1 solutions"
author: "Amer Abuhasan"
desription: "Reading in data and creating a post"
date: "09/21/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
railroad <- read_csv("_data/railroad_2012_clean_county.csv")
railroad


```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data
The data in the data set explained is the state,county, and total employments. I filtered them out to there individual categories. each one has a set of information based on his list, it includes means,medians,mins, and max's.

```{r}
#| label: summary
summary(railroad)
# number of states 
railroad%>%
  select(state)%>%
  n_distinct(.)
# number of countys
railroad%>%
  select(county)%>%
  n_distinct(.)
# number of total employees
railroad%>%
  select(total_employees)%>%
  n_distinct(.)




```
