---
title: "Challenge 2"
author: "Lai Wei"
desription: "Data wrangling: using group() and summarise()"
date: "10/26/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - faostat
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(summarytools)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

-   railroad\*.csv or StateCounty2012.xls ⭐
-   FAOstat\*.csv or birds.csv ⭐⭐⭐
-   hotel_bookings.csv ⭐⭐⭐⭐

```{r}
hotel <- read.csv("_data/hotel_bookings.csv")
```

Read the hotel_bookings data as hotel. 

## Describe the data

```{r}
hotel %>% 
  select(reservation_status) %>% 
  distinct()
```
There are 3 different reservation status of the hotel, 'check-out', 'canceled' and 'no-show'.

## Provide Grouped Summary Statistics

Data summary
```{r}
hotel %>% 
  summarise(across(c(country,market_segment),n_distinct))
``` 
Count the number of unique 178 countries and 8 different market_segments of hotel data

Group_by() application
```{r}
hotel %>% 
  group_by(reservation_status) %>% 
  summarise(num_adult = sum(adults,na.rm = TRUE),
            num_child = sum(children,na.rm = TRUE),
            num_baby = sum(babies, na.rm = TRUE))
```
In total, we can see how many adults, children and babies were checking in successfully. The information may help the hotel to arrange more facilities 
for children or babies. 

Data summary visualization 
```{r}
print(summarytools::dfSummary(hotel,
                        varnumbers = FALSE,
                        plain.ascii  = FALSE, 
                        style        = "grid", 
                        graph.magnif = 0.70, 
                        valid.col    = FALSE),
      method = 'render',
      table.classes = 'table-condensed')
```

### Explain and Interpret

I choose this data is because I did the middle level in summer session and this time I want to practice something which is more complex and fun for increasing my R skill. 
