---
title: "Challenge 2 Erika Nagai"
author: "Erika Nagai"
desription: "Data wrangling: using group() and summarise()"
date: "09/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - faostat
  - hotel_bookings
---

First of all, let's install all the packages convenient for this challenge.

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(summarytools)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xls ⭐
-   FAOstat\*.csv or birds.csv ⭐⭐⭐
-   hotel_bookings.csv ⭐⭐⭐⭐

I decided to analyze "hotel_booking.csv data".

```{r}
# read the data as a tibble
data <- read_csv("_data/hotel_bookings.csv")
data
```

## Describe the data

Let's explore the data.

```{r}
view(dfSummary(data))
```

From this summary, we can see that this data is about reservations made for two types of hotel (City Hotel, Resort Hotel) from 2015 until 2017.

The data includes 1. Status of reservation (cancelled, checkout, no-show) 2. The length of stay (arrival date, days in total, days during weekdays and weekend), 3. Guest information (# of adults, children, babies, repeater or not, previous cancellation history) 4. Reserved service type (room type, meal, car parking, special request) 5. Others (used agency, distribution channel etc)

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

Let's explore the differences in reservation trends by the type of hotel (City Hotel vs. Resort Hotel)

!! Questions: The below code shows the statistic data of only Resort Hotel.

```{r}
data %>% 
  group_by(hotel) %>%
  descr(stats = c("mean", "sd", "min", "med", "max"))

```

Make a column of total nights of stay

```{r}
data <- data %>% 
  mutate(total_nights = stays_in_week_nights + stays_in_weekend_nights)

```

Customer and Stay Trends by Hotel Type

```{r}
data %>%
  group_by(hotel) %>%
  summarise(
    ave_adults = mean(adults, na.rm = TRUE),
    ave_children = mean(children, na.rm = TRUE),
    ave_babies = mean(babies, na.rm = TRUE),
    avg_total_nights = mean(total_nights, nr.rm = TRUE))


```

Guest who stay at Resort hotel tend to stay longer than those who stay at Resort hotel. Guests at both hotels are less likely to be accompanied by children, but resort hotels have more people with children and babies than city hotels!

### Explain and Interpret

I will compare **the group with children (at least one child or baby) to the group without children**. I chose this group because I thought that the presence of children on a trip significantly impacts the stay trend and that hotel companies would want to know what the groups with children want.

```{r}
# First classify if the group is with at least one kid/baby.

data <- data %>% 
  mutate(Kids_total = children + babies) %>%
  mutate(Kids = case_when(
    Kids_total == 0 ~ "No Kid",
    Kids_total >= 1 ~ "With Kid"
  ))

data
```

Now, let's find out the differences in trends between groups with and without children.

```{r}
data %>%
  group_by(Kids) %>%
  summarise(
    avg_total_nights = mean(total_nights, nr.rm = TRUE),
    repeated_guest = mean(is_repeated_guest, nr.rm = TRUE),
    lead_time = mean(lead_time, nr.rm = TRUE)
  )
```

Here are some findings

-   Groups with kids tend to stay a little longer than groups without kids.

-   Groups without kids are likely to book much earlier than groups with kids.
