---
title: "Challenge 1 Kristin Abijaoude"
author: "Kristin Abijaoude"
desription: "Reading in data and creating a post"
date: "09/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - kristin_abijaoude
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
#| label: read
railroads<- read_csv("_data/railroad_2012_clean_county.csv")
railroads
```

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: Summary
summary(railroads)
```

In total, we have 2930 entries in this dataset, with 2930 counties across the US accounted for. Some counties have a minimum of 1 railroad employee, while Cook County, IL has the most with 8207 railroad employees.

Here are the first 6 rows of the dataset. In total, there are 2930 rows and 3 columns: state, county, and total employees.

```{r}
head(railroads)
```

```{r}
dim(railroads)
```

```{r}
colnames(railroads)
```

Let's narrow it down to Massachusetts, where we are now. There are 12 counties in Massachusetts, with Barnstable County with the least amount of railroad employees with 44 of them, while Middlesex County has the most with 673 railroad employees.

```{r}
filter(railroads, state == "MA")
```

```{r}
filter(railroads, `total_employees` > 1000)
```

There are 27 counties with more than 1000 total railroad employees, while there are 488 counties with fewer than 5 (yes, you read that correctly) railroad employees.

```{r}
filter(railroads, `total_employees` < 5)
```

```{r}
filter(railroads, `total_employees` > 1000 & `state` == "CA")
```

When I used the filter code to only show California, the nation's most populous state, it's unsurprising that there would be a lot of railroad employees.
