---
title: "Challenge-3"
author: "Said Arslan"
desription: "Tidy Data: Pivoting"
date: "10/01/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_3
  - eggs

---

```{r}


library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

```


## Read in data


```{r}

eggs <- read.csv("_data/eggs_tidy.csv")

```


### Briefly describe the data


```{r}

dim(eggs)
head(eggs)
tail(eggs)
sample_n(eggs, 10)

unique(eggs$year)

```

The dataset includes monthly prices of 4 types of boxes of eggs (in cents) beginning from January 2004 to December 2013.


The column variables `large_half_dozen`, `large_dozen`, `extra_large_half_dozen` and `extra_large_half_dozen` are actually not variable names but type and number of eggs in boxes. On the other hand, the values under these 4 columns show prices. Thus we should reorganize these columns so that each row represents an observation. Column names in tidied version are: `month`, `year`, `type` and `price`. 



## Anticipate the End Result


originally our dataset has 120 rows and 6 columns. I will use 4 variables for pivoting. So 2 variables will be used as identifiers for each observation. Therefore I should have 480(120*4) rows and 4(2+2) columns in the longer format.



## Pivot the Data


```{r}

eggs.pivoted <-pivot_longer(eggs, cols= ends_with("dozen"),
                 names_to= "type",
                 values_to= "price")

head(eggs.pivoted)

```


Also, it would be better if the first column is `year` and the second column is `month` because main identifier for time of an observation is `year`. Values of `month` are repetitive.


```{r}


eggs.pivoted <- eggs.pivoted[, c("year", "month", "type", "price")]

head(eggs.pivoted)

```


To check if my calculation for dimension of new dataset is correct or not, let me look at row and column numbers of pivoted dataframe.


```{r}

cat("Number of rows are: \n")
nrow(eggs.pivoted)

cat("Number of columns are: \n")
ncol(eggs.pivoted)

```

In new dataset, each row shows price of a specific type of box of eggs in a month of a year.
Now, using this dataframe, we can do further price analysis by grouping based on type.
