---
title: "Challenge 4"
author: "Mariia Dubyk"
desription: "More data wrangling: pivoting"
date: "11/03/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
  - abc_poll
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:


2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐

```{r}
abc_poll<-read.csv("_data/abc_poll_2021.csv")
library(summarytools)
view(dfSummary(abc_poll))
```

### Briefly describe the data
The dataframe shows responses to the questionnaire. The respondents are adults (age 18-91) from the USA. We can see some demographic data like education, income, age, state, etc (column 4-17). Responses to other questions (probably about views or experience) are in columns 18-28. Column 2 probably identifies language of the respondent. Dataset contains of 527 observations and 30 variables (not including "id"). Some variables are dichotomous, and other variables may have several possible values.  

## Tidy Data (as needed)

From the first sight data looks tidy. 
- We have each response in a separate column.
- I did not find missing data. 
- I need to look closer at column 8 "ppethm" and think if it would be proper to make two variables of it or remove information after comma.
- Also it might be easier to understand data if columns had different names. 
```{r}
abc_poll<-rename(abc_poll, language = xspanish, age =  ppage, education5 = ppeduc5, education = ppeducat, gender = ppgender, ethnicity = ppethm, household_size = pphhsize, income = ppinc7, marital_status = ppmarit5, region = ppreg4, rent = pprent, state = ppstaten, work = PPWORKA, employment = ppemploy)
abc_poll <- abc_poll%>%
  mutate(ethnicity = str_remove (ethnicity, ", Non-Hispanic"))

```

Any additional comments?

## Identify variables that need to be mutated
Some variables should turned into factors. I will try to do it with QPID. This variable need to be reordered so that for example bar chart looked better.

Document your work here.

```{r}
abc_poll<-abc_poll %>%
  mutate(QPID = fct_recode(QPID, "dem" = "A Democrat",
                                "rep" = "A Republican",
                                "ind" = "An Independent",
                                "na" = "Skipped",
                              "other" = "Something else")) %>%
  mutate(QPID = fct_relevel(QPID, "dem", "ind", "rep","other", "na"))

ggplot(abc_poll, aes(QPID)) + geom_bar()

```

Any additional comments?
