---
title: "Challenge 1 - Darron Bunt"
author: "Darron Bunt"
desription: "Reading in data and creating a post"
date: "10/09/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - birds
  - darron bunt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Step 1 - Read in the Data

*Read in one (or more) of the following data sets, using the correct R package and command.*

-   birds.csv ⭐⭐

```{r}
birds <- read_csv("_data/birds.csv")
```

*Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.*

## Step 2 - Describe the data

*Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).*

Ok, so from what's been read in above, we know that the birds dataset has 30,977 rows and 14 columns. Eight of those columns are character-based, while the remaining six are number-based. Neat.

So now if I run birds, I should get a tibble, and in theory that tibble is going to help me perform a high-level description of the data. 

```{r}
birds
```
The data appears to show the worldwide data on the historical value of 1,000 head of five different birds. Specifically, the dataset includes information relating to chickens, ducks, geeese and guinea fowls, pigeons and other birds, and turkeys, from 601 areas of the world (some countries, some regions), dating from 1961 to 2018.     

```{r}
count(birds, Item)
count(birds, Area)
count(birds,Year)
```
Judging by the flag descriptions, this data has come from a variety of sources, most commonly FAO (Food and Agriculture Organization) estimates and official data. 

```{r}
count(birds,`Flag Description`)
```
Several columns contain repetitive data; the value for Domain Code and Domain is the same across all entries in the dataset (QA for the former; Live Animals for the latter), as is the value for Element Code and Element (5112 and Stocks, respecitvely). The columns for Year Code and Year repeat the same data. The Unit is also the same for the entire dataset (1,000 head). 

I used a variety of count commands to ascertain the above; for reference I have included that for Domain Code and Domain.

```{r}
count(birds,`Domain Code`)
count(birds, Domain)
```

```{r}
#| label: summary

```
