---
title: "Challenge 4 Instructions"
author: "Kimberly Darkenwald"
desription: "More data wrangling: pivoting"
date: "10/24/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
  - abc_poll
  - eggs
  - fed_rates
  - hotel_bookings
  - debt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.xlsx or organiceggpoultry.xls⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions.xlsx ⭐⭐⭐⭐⭐

```{r}
abcpoll2021 <- read_csv("_data/abc_poll_2021.csv")
abcpoll2021
dim(abcpoll2021)
colnames(abcpoll2021)



abcpoll2021 %>%
select("ppeducat", "QPID", "ABCAGE") %>%
  group_by("ppeducat" =="A Democrat") %>%
  summarize(mean)
  
```

### Briefly describe the data

I would like to know the average age in relation to political identification as well as their education level.  



```{r}
library(tidyr)

table(abcpoll2021$ppethm)

abc_poll<-abcpoll2021%>%
  mutate(ethnic = str_remove(ppethm, ", Non-Hispanic"))%>%
  select(-ppethm)

table(abc_poll$ethnic)


```
```{r}
table(abcpoll2021$ppinc7)


abc_poll <- abcpoll2021 %>%
  mutate(income= case_when(ppinc7=="Less than $10,000" ~ 1,
                         ppinc7=="$10,000 to $24,999" ~ 2,
                         ppinc7=="$25,000 to $49,999" ~ 3,
                         ppinc7=="$50,000 to $74,999" ~ 4,
                         ppinc7=="$75,000 to $99,999" ~ 5,
                         ppinc7=="$100,000 to $149,999" ~ 6,
                         ppinc7=="$150,000 or more" ~ 7))%>%
  select(-ppinc7)

table(abc_poll$income)



```

## Identify variables that need to be mutated

I mutated two variables: (1) the ethnicity column and (2) The income column.



