---
title: "Challenge 1"
author: "Shriya Sehgal"
desription: "Reading in data and creating a post"
date: "11/1/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data





```{r}
library(readxl)
dataset<-read_excel("_data/wild_bird_data.xlsx", skip=1)
dataset

```

## Display the dimensions of the data
```{r}
dim(dataset)

```

The data presented in the dataset is in the form of an excel and stored in the variable dataset. The first row which consists of the description of the dataset has been ignored in our analyses. The dataset consists of 2 columns ('body weight' and 'population_size') with the dimensions [146, 2].


## Describe the data



```{r}
# Checking if there are any null values.
is.null(dataset)

```
```{r}
#Summary of the dataset
summary(dataset)
```



The dataset consists of 2 columns body_weight and population_size containing the numerical values. Before our analysis, we need to make sure that there is no NULL values present in the data. We then use the summary(dataset) that summarizes the statistical data of out dataset and provides the mean, median, min and max values. 
One can see that the birds which are heavier and have higher body weight are higher in number clearly stating that the survival rate of the bird is directly proportional to the body weight of the bird.
