---
title: "Challenge 1 "
author: "Emma Narkewicz"
description: "Emma Narkewicz completes Challenge 1"
date: "09/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - emma_narkewicz
  
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information
    (e.g., tables, etc)

## Read in the Data

I read in the railroad_2012_clean_county.csv⭐ Using the read_csv
function.

```{r}
#First I installed the readr function
install.packages("readr")
#Next I loaded the tidyverse library
library(tidyverse)
#open data set
Trains <- read_csv("_data/railroad_2012_clean_county.csv")
Trains


```

## Describe the data

I chose to analyze the railroad_2012_clean_county.csv data set as a
beginner. There are 3 columns in this data set and 2930 rows. The
columns values are "state", "county" and "total_employees". The state
column provides the 2 letter abbreviation for the state in which the
railroad data was collected from in 2012.

There were 53 "states" included in this data set, which was 50 states,
as well as the District of Columbia, the Armed Forces, and the Armed
Forces Pacific. The county column refers to every county in the state
that railroad data was collected from, with data from 2930 counties
included in the data set. The total_employees column refers to the total
number of railroad employees in said county in the year 2012. Data was
only collected in the year 2012, meaning that this data set does not
show trends over time.

```{r}
#Determining min, max, median, and mean number of total employees

summarise(Trains, min = min(total_employees), max = max(total_employees), median = median(total_employees), mean = mean(total_employees))


```

There was a wide range of total county railroad employees in counties in
2012, with a minimum of 1 employee and a maximum of 8207 employees. The
median number of employees was 21 and the mean number of employees was
\~87. A few counties with high totals employees in some counties skewed
the mean higher than the median.

```{r}
#Creating a boxplot of total county railroad employees
ggplot(Trains, aes(state,total_employees)) + geom_boxplot() + theme_minimal() + labs(title = "Total County Railroad Employees in 2012")


#Finding IQR of total employees
summarise(Trains, IQR= IQR(total_employees))

#removing outlier from Trains
Outlier_free_trains <- Trains %>%
  filter(total_employees<8200) %>%
  arrange(total_employees)

Outlier_free_trains


#Calculating summary statistics with outlier removed
summarise(Outlier_free_trains, outlier_free_min = min(total_employees), outlier_free_max = max(total_employees), outlier_free_median = median(total_employees), outlier_free_mean = mean(total_employees))


```

Looking at the above box-plot, the maximum of 8027 total employees looks
like a visual outlier. To confirm, I calculated the IQR was 58
employees. From running summary statistics, I know that Q1 is 7
employees and Q3 is 65 employees. Using the rule that Q3 + 1.5(IQR) is
the threshold for upper outliers, which equals 152 employees, the
maximum value of 8,207 employees meets the criteria for outlier.
However, removing this point from the data would not change the median
from 21 employees and would only decrease the mean total employees
slightly from \~87 to \~84.4 employees.

```{r}
#Identifying 10 counties with most total employees
Trains %>%
  arrange(desc(total_employees)) %>%
  slice(1:10)

```

After seeing visually from the box plot that most counties had lower
numbers of total employees, I used the arrange() and slice() functions
to identify the counties with the top 10 total railroad employees in
2012. These counties were in the states of IL,TX,NE,NY,VA,FL, and CA.

```{r}
  
#Recode `total_employees` using case_when, numerical > categorical
Trains <- Trains %>%
  mutate(Number_employees = case_when(
    total_employees>= 5000 ~ "5000+ employees",
    total_employees>= 2500 & total_employees<5000 ~ "2500-4999 employees",
    total_employees>= 1000 & total_employees<2500 ~ "1000-2499 employees",
    total_employees>= 500 & total_employees<1000 ~ "500-999 employees",
    total_employees>= 100 & total_employees<500 ~ "100-499 employees",
    total_employees <100 ~ "0-99 employees")
  )

#Create table with the recoded values
table(select(Trains, Number_employees))

#Create prop.table with recoded values
prop.table(table(select(Trains, Number_employees)))


```

Lastly, I re-coded the numerical total_employees to the categorical
number_employees mutate() and case_when() and created a table and
prop.table with these new categorical values. Just as the box_plot
suggested, the majority of counties in the data set had fewer than 100
railroad employees in 2012, nearly 82%. Additionally over 95% of
counties had less than 500 total employees, and only 9 counties total
had more than 2500 total employees.
