---
title: "Challenge 1 "
author: "Niyati Sharma"
desription: "Reading in data and creating a post"
date: "10/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
birds_df <- read_csv("_data/birds.csv")
birds_df

```



## Description 
Birds dataset has 30,977 rows and 14 columns. Eight of those columns are character-based, while the remaining six are integer based. Through Summary function we can calculate the duration of the dataset(year 1961-2018) and there are 1036 cells where the value is NA.From count function we can say that there are 5 types of birds in the dataset.


```{r}
summary(birds)

```
```{r}
count(birds, Item)
```

