---
title: "Challenge 1"
author: "Hezzie Phillips"
description: "Reading in data and creating a post"
date: "09/21/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Overview

Below please find my overview of the railroad dataset from 2012:  

-   railroad_2012_clean_county.csv ⭐

## Code to read data
```{r}
railroadclean<-read_csv("_data/railroad_2012_clean_county.csv")

```


## Summary

We see that there are three variables in this table: state, county and total employees.  

The data seems to be tabulating the number of railroad employees in each county.  

We can see that the largest employment number in a county is 8207 (in Cook County Illinois) and the smallest is one (which happens in 145 counties). With such a wide range it is useful, perhaps, to look at the median which is 21 people.

```{r}
#| label: summary
summary(railroadclean)
distinct(railroadclean)

```
