---
title: "Challenge 6 - Abby Balint"
author: "Abby Balint"
description: "Visualizing Time and Relationships"
date: "10/25/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_6
  - abby_balint
  - fed_rate
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(lubridate)
library(hrbrthemes)
library(treemap)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


### Briefly describe the data

Reading in the Fed Fund Rate dataset.

This data is likely government data for the federal funds rate which determines the rate that banks borrow money from each other at. This data set contains to variables ranging from date, target/upper target/lower target rates, actual rates, as well as GDP, unemployment, and inflation rates. The dataset contains 904 rows initially, beginning in July of 1954 and ending in March of 2017. The dataset is not complete as several of the target rates and upper/lower target rates are not included. This may be because they were not reported during earlier years, or standardly reported every month in general.

My goal here is to look at factors like unemployment and inflation since the year 2000 as well as generate some charts for analysis.

```{r}
rates <- read_csv("_data/FedFundsRate.csv")
head(rates,2)
```


## Tidying the Data
This data set is already pretty tidy since the rates are in a reportable format, however I wanted to use this opportunity to practice mutating month/day/year columns into a date. I used the make_date lubridate function to create a new variable called Date. I filtered to the year 2000 or later for my graphing purposes. I also filtered out some blank values from some of the values I was using to chart so that my line graphs wouldn't have any breaks. There isn't much to pivot/mutate/recode here as most of these variables are just straightforward numerical.

```{r}

rates_tidy <- rates %>%
  mutate(Date = make_date(`Year`, `Month`, `Day`)) %>%
  mutate(`Employment Rate` = (100 - `Unemployment Rate`)) %>%
  filter(`Year` >= 2000) %>%
    drop_na(`Inflation Rate`) %>%
  drop_na(`Unemployment Rate`) %>%
  drop_na(`Real GDP (Percent Change)`)
summary(rates_tidy)
```

```{r}

```

## Time Dependent Visualization

In this visualization, I generated a line graph for the inflation rate since 2000. I used the date variable I generated above to create this chart. I also implemented a color scheme and played around with the line thickness and type.

```{r}
rates_tidy %>%
  ggplot(aes(x=`Date`, y=`Inflation Rate`)) +
  geom_line(color="#00658E", size=1, alpha=6, linetype=1) +
  ggtitle("Change in Inflation Rate Since Year 2000") +
  theme(panel.background = element_rect(fill="lightblue"))

```


## Visualizing Part-Whole Relationships

In this visualization, I was looking for something that I could create a part-whole chart for. Since there is not many categorical variables in this dataset, I used year. The below tree map makes it easy to see which years had some of the highest unemployment rates since the year 2000.

```{r}
rates_tidy %>%
  treemap(index=c("Year"), vSize="Unemployment Rate", title="Unemployment Rate Comparison, 2000-2017")
```

