---
title: "Homework 1"
author: "Julian Castoro"
desription: "Challenge 1 submission on wild bird data"
date: "08/02/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - challenge1
  - Julian Castoro
  - wild_bird_data
  - 
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(readxl)
library(ggplot2)
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE)

old<-options(pillar.sigfig = 2)

```


### With only skipping the first line here is what wild_bird_data.xlsx looks like
```{r}
birdsData <- read_excel("_data/wild_bird_data.xlsx",skip=1)
birdsData%>%
  head()%>%
  arrange(`Wet body weight [g]`)
```

While this data source is lacking the names of each bird, it appears the 
wet body weight in grams was collected for a variety of birds(146). This weight is
then tied to an estimated population size, I say estimated because the numbers 
are not integers.

Wet body weight[g] reflects the weight of a particular bird while alive as well
as that birds (est?) population size.

```{r}
count(birdsData)
```
## Body weight stats
Smallest bird:
```{r}
smallestBird<-birdsData%>%
  slice(1)

smallestBird
```

Largest bird:
```{r}
birdsData%>%
  tail(n=1)
```
To give some perspective: 

  Robin average weight: 70g.\n

  Pelican average weight: 11,000g

### Average
```{r}
birdsData%>%
  summarise('average weight'=mean(`Wet body weight [g]`))
```

## population size vs body weight plot
Here I wanted to show how body weight correlates with population size. The
raw chart had some outliers which led me to focus in on birds with a 
body weight of less than 2.5kgs

### raw graph
```{r}
#| message: false
graph<-
  birdsData%>%
  ggplot(aes(`Wet body weight [g]`,`Population size`)) + geom_point() + geom_smooth(method="lm")

graph
```
### With constraints
We see once we focus in on smaller birds, weight< 500g, a bit more of a clear
trend in pop size vs weight
```{r}
#| warning: false
#| message: false
graph + xlim(c(0, 500)) 
```

# Conclusions

This data could be used to draw conclusions about the populations of birds based
on their wet body mass. 

