---
title: "Challenge-1"
author: "Said Arslan"
desription: "Reading in data and creating a post"
date: "09/25/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - birds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

```

I will analyze the `birds.csv` dataset in this challenge.

## Read Data

```{r read_data}

birds <- read.csv("_data/birds.csv")

```

`birds` variable is created to internalize the dataset so I can analyze and manipulate it.

## Describe the data


```{r about-data-1}

dim(birds)
head(birds)
tail(birds)

```


In the dataset, there are 14 variables (columns) and 30977 observations (rows). From head and tail tables, roughly speaking, each observation includes stock level of a specific type of animal in a country in a year.



```{r about-data-2}

sapply(birds, function(x) n_distinct(x)) # number of unique values under each column

table(birds$Item) # types of birds

table(birds$Flag) # flag types

country.names <- unique(birds$Area) # countries included in the dataset
print(country.names)

years.of.observations <- range(birds$Year) # range of data in terms of years
years.of.observations


```
If we assess the number of unique values in each column with the names of variables, we can say that this dataset is extracted from a large stock dataset because it includes only live animals as `Domain` and only 5 types of domestic fowl is included under `Item` variable which are chicken, duck, geese, turkey and other types of birds. Besides that, only one type of element (stocks) is included in the data.

Observations are from 248 different countries and regions. Observations are years from 1961 to 2018. The dataset covers almost every region and countries of the World.

The core of each observation is given under `Value` variable, which should be number of stocks in thousands. All of other variables are basically contain characteristic information about observations.

Data might be obtained from poultry farms in each country from 1961 to 2018. It includes aggregated data.


## Missing Data

```{r missing-data}

sum(is.na(birds$Value))

sum(is.null(birds$Value))

```

It looks 1036 value are missing/ not observed in the dataset.


## Stock levels by country and by bird type

```{r stock-levels-by-country-and-region}

birds %>% 
  select(Area, Value) %>% 
  group_by(Area) %>% 
  summarize(Total = sum(Value)) %>% 
  arrange(desc(Total))

```

Summarize table shows the total of stock levels in 58 years for each predefined region and country. Among countries, China had the largest stock of poultry over 58 years.


```{r stock-levels-for-each-type-of-birds}

stock.levels <-  birds %>% 
  select(Item, Value) %>% 
  group_by(Item) %>% 
  summarize(Total = sum(Value, na.rm = TRUE)) %>% 
  arrange(desc(Total)) %>% 
  mutate(Percentage = 100*round(proportions(Total), digits=3))
  

stock.levels


```

In terms of bird types, more than 90% of poultry stocks consisted of chickens, 5% ducks and so on.