---
title: "Challenge 1"
author: "Mariia Dubyk"
desription: "Reading in data and creating a post"
date: "09/22/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in the Data

```{r}
library(readr)
birds<-read_csv("_data/birds.csv")
```

## Describe the data
Data set represents existence of certain species of birds in different geographic areas (countries, continents, world in general) from 1961 to 2018. We have observation of number of certain birds in population each year (from 1961 to 2018). It is difficult to indicate cases. From columns "Unit" and "Value" we understand number of birds. The variables are for example "Year" and "Area". So with the data set we may look how situation changed in some area, focus on comparison of populations of different species in one area, etc. The information was probably gathered from the farmers or farming businesses in certain regions.

```{r}
birds
subset (birds, select = c("Domain", "Area", "Element", "Item", "Year", "Unit", "Value", "Flag", "Flag.Description"))
birds2 <- birds [,c("Domain", "Area", "Element", "Item", "Year", "Unit", "Value", "Flag", "Flag.Description")]
birds2


```
```{r}
summary (birds2)

```

```{r}
str(birds2)

```
