---
title: "Challenge 3"
author: "Mariia Dubyk"
desription: "Tidy Data: Pivoting"
date: "11/01/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_3
  - animal_weights
  - eggs
  - australian_marriage
  - usa_households
  - sce_labor
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(readxl)
marriage <- read_excel("_data/australian_marriage_law_postal_survey_2017_-_response_final.xls",
                       sheet = "Table 2",
                       col_names = c("Division", "Yes", "Yes_percentage", "No", "No_percentage", "Total_clear", "Total_clear_percentage", "Delete", "Delete", "Delete", "Response_not_clear", "Response_not_clear_percentage", "Non_responding", "Non_responding_percentage", "Total", "Total_percentage"))

marriage <- slice(marriage, 8:180)
marriage <- select(marriage, "Division", "Yes", "Yes_percentage", "No", "No_percentage", "Total_clear", "Total_clear_percentage", "Response_not_clear", "Response_not_clear_percentage", "Non_responding", "Non_responding_percentage", "Total", "Total_percentage")

```

### Briefly describe the data

The dataset shows responses to the question in Australian Marriage Law Postal Survey 2017. We can see data gathered in different divisions of States and Territories in Australia. The responses are "Yes", "No", "Not clear" and "No response". The table also shows percentage of each answer and total numbers.

To tidy data I will leave only divisions in the first column. I am going to move State and Territories to another column by adding a variable (as it was shown during the class).I will also remove percentages and total numbers from the table.

```{r}

marriage <- marriage%>%
  mutate(StateTerritory = case_when(str_ends(Division, "Divisions") ~ Division), NA_character_)%>%
  fill(StateTerritory, .direction = "down")
marriage <- filter(marriage, !str_detect(Division, "Division|Australia"))
marriage <- select(marriage, "Division", "Yes", "No", "Response_not_clear", "Non_responding", "StateTerritory")
marriage
```


## Anticipate the End Result

In the final dataframe we will have 4 rows (with answers "Yes", "No","Response_not_clear", "Non_responding") for each division. 


### Example: find current and future data dimensions

Lets see if this works with a simple example.

```{r}
#| tbl-cap: Example

df<-tibble(country = rep(c("Mexico", "USA", "France"),2),
           year = rep(c(1980,1990), 3), 
           trade = rep(c("NAFTA", "NAFTA", "EU"),2),
           outgoing = rnorm(6, mean=1000, sd=500),
           incoming = rlogis(6, location=1000, 
                             scale = 400))
df

#existing rows/cases
nrow(df)

#existing columns/cases
ncol(df)

#expected rows/cases
nrow(df) * (ncol(df)-3)

# expected columns 
3 + 2
```

Or simple example has $n = 6$ rows and $k - 3 = 2$ variables being pivoted, so we expect a new dataframe to have $n * 2 = 12$ rows x $3 + 2 = 5$ columns.

### Challenge: Describe the final dimensions


```{r}


```

Any additional comments?

## Pivot the Data

Now we will pivot the data, and compare our pivoted data dimensions to the dimensions calculated above as a "sanity" check.

### Example

```{r}
#| tbl-cap: Pivoted Example

df<-pivot_longer(df, col = c(outgoing, incoming),
                 names_to="trade_direction",
                 values_to = "trade_value")
df
```

Yes, once it is pivoted long, our resulting data are $12x5$ - exactly what we expected!

### Challenge: Pivot the Chosen Data

After pivoting the case is number of certain response in a division. I think it may be easier to visualize the data using this type of table. So we may say that data is tidy because it is easier to make next steps in analyzing data. 

```{r}

marriage_new<-pivot_longer(marriage, col = c(Yes, No, Response_not_clear, Non_responding),
                 names_to="Response_type",
                 values_to = "Number")
marriage_new
```

