---
title: "Challenge 2"
author: "Matthew O'Neill"
desription: "Data wrangling: using group() and summarise()"
date: "10/05/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - faostat
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xls ⭐
-   FAOstat\*.csv or birds.csv ⭐⭐⭐
-   hotel_bookings.csv ⭐⭐⭐⭐

```{r}
data <- read_csv("../posts/_data/birds.csv")
```


## Describe the data

Below is the first few rows of the dataset. It is very similar to the FAO dairy dataset, which I looked through in Challenge 1. At first glance, thess data appear to be a lot simpler than the dairy data, as we may just be dealing with the count of chickens accross fars in different Countries over time. But there may be more bird types that we don't see right away. The data was once again likely gather by the Food and Agriculture Organization in the US. 

```{r}
#| label: summary

head(data)

```
Below is a breakdown of item types in our dataset, and it is clear we are workign with more than just Chickens. The data also includes Turkeys, Ducks, Geese, and Pigeons. It's likely that different countries will produce differnt proportions of each.

```{r}
#| label: Item

item <- select(data, "Item")
table(item)

```


## Grouped Summary Statistics

Below we can see two tables, one showing the mean number of each kind of bird in each area. The second sorts these numbers in order to show that the largest count of birds in any one country is Chickens with 1.5 Million in total.

```{r}

bird_type <- data %>%
  group_by(`Area`,`Item`) %>%
  replace_na(list(`Value` = 0))%>%
  summarise(mean(`Value`))

bird_type
bird_type %>%
  arrange(desc(`mean(Value)`))
  

```
Seeing as the United States seems to be one of the largest producers of birds for meat, let's take a closer look at their production.

```{r}

data %>%
  filter(`Area`=="United States of America")%>%
  select(`Area`,`Item`,`Year Code`,`Value`)%>%
  group_by(`Item`)%>%
  summarize(mean = mean(`Value`),median = median(`Value`), std = sd(`Value`), max = max(`Value`), min = min(`Value`))

  

```
### Explain and Interpret

It appears that the US produces mostly Chicken, with a mean and median production over the years of around 1.3 million chickens. They produced far fewer Turkeys over the same time period, with an average of 206K per year, and barely any Ducks in comparison to the previous two subgroups. We also see that the US did not produce any Geese or Pigeons as food.

Comparing maximum and minimum values over time, it's interesting to point out that the best year for Turkey production was still less than half that of the worst year for Chicken production. This makes some sense knowing US culture, as Turkey is a much more seasonal meat than Chicken.

Finally it's important to point out that there was quite a bit of variance in Turkey production, with a standard deviation of nearly 70K per year despite it's mean. 
