---
title: "Challenge 1"
author: "Siddharth Nammara Kalyana Raman"
desription: "Reading in data and creating a post"
date: "11/27/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
library(readxl)
wild_birdData<-read_excel("_data/wild_bird_data.xlsx",skip=1)
wild_birdData
```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.




## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

##Dimensions and Data Extraction

```{r}
dim(wild_birdData)
```

The given data is in the form of excel and is read using the readxl library. The data is stored in wild_birdData. The first row of the actual dataset shows reference and from where the data is being extracted which is not relevant for our study. Hence, the first row is skipped. The dataset consists of two columns 'bodyweight' and 'population_size'. The dimensions of the dataset is [146,2].

#null check

```{r}
is.null(wild_birdData)
```

```{r}
#| label: summary
summary(wild_birdData)

```

The wild birds dataset consists of two columns or attributes 'Wet body Weight' and 'Population Size'. We have checked that there are no null values in the dataset. If we have null values in the data set we need to preprocess it before going to further evaluations. Moreover, the birds with less body weight are low in number and the birds that are heavier are greater in number. From this we can infer that the population of the birds are directly proportional to their body weight. All the other statistical information about the dataset is shown above.