---
title: "Challenge 5"
author: "Jack Sniezek"
description: "Introduction to Visualization"
date: "12/06/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_5
  - cereal

---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  create at least two univariate visualizations
   - try to make them "publication" ready
   - Explain why you choose the specific graph type
5)  Create at least one bivariate visualization
   - try to make them "publication" ready
   - Explain why you choose the specific graph type

[R Graph Gallery](https://r-graph-gallery.com/) is a good starting point for thinking about what information is conveyed in standard graph types, and includes example R code.

(be sure to only include the category tags for the data you use!)

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   cereal.csv ⭐


```{r}

cereal <- read_csv("_data/cereal.csv")
cereal
summary(cereal)

```

## Briefly describe the data

The cereal dataset is made up of 4 variables: Cereal, Sodium content, Sugar content, and Type(Adult or Child). There are 20 cereal types.

## Tidy Data (as needed)

The data appears to be tidy already, there is no missing data.

## Univariate Visualizations

My first visualization is to put the sugar levels of Children's cereals into a bar graph. I felt like a bar graph was the best way to view which cereals had the most/least sugar as well as which ones were similar.

The second visualization I did was to use a bar graph showing the sodium levels of all cereals. Again, I thought this was the best type of graph to show visually which cereals were similar and different.
```{r}

cereal %>%
  filter(Type == "C") %>% 
  arrange(Sugar) %>%   
  mutate(Cereal=factor(Cereal, levels=Cereal)) %>%
ggplot(aes(x=Cereal, y=Sugar)) + 
  geom_bar(stat = "identity", fill = "darkorchid") +
  theme_minimal() +
  coord_flip() +
    ggtitle("Sugar Level of Children's Cereals")



cereal %>%
  mutate(Cereal = factor(Cereal))%>%
  ggplot(aes(x=Cereal, y = Sugar)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  theme_light() +
  coord_flip() +
  ggtitle("Sodium Levels of Cereals") +
  labs(y = "Sugar Content", x = "Sodium Content")
```

## Bivariate Visualization(s)

My Bivariate visualization was comparing the sugar content of cereals to the sodium content of cereals. I did this by using a scatterplot to represent the data points with Sugar as the y axis and Sodium as the x axis. I felt like this was the best way to compare cereals and view the similarities and differences in each cereal.

```{r}
cereal%>%
  ggplot(aes(x = Sugar,y = Sodium)) +
  geom_point(color = "green") +
  ggtitle("Sugar Compared to Sodium Levels in Cereals") +
  labs(y = "Sugar Content", x = "Sodium Content")

```

Additional Thoughts:

1. When I work on the next challenges I want to think about labeling the data within the chart, such as the scatterplot points and maybe even the bars in a bar graph if it is fitting.

2. I want to think about making data points different colors and adding a legend in the next challenges

3. Surprised that Raisin Bran is the worst cereal for both categories, its not even a great tasting cereal
