---
title: "Challenge 4"
author: "Guanhua Tan"
desription: "More data wrangling: pivoting"
date: "10/1/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - eggs
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.xlsx or organiceggpoultry.xls⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions.xlsx ⭐⭐⭐⭐⭐

```{r}
library(dplyr)
library(readxl) 
poultry<-read_excel("_data/poultry_tidy.xlsx")
poultry

```
### Briefly describe the data
The dataset records the prices of Chicken and its different parts from Jan. 2004 to Jan 2013. The column names includes Product, Year, Month, Price_Dollar. There are a few values (price) missing.

## Tidy Data (as needed)

First, I want data have a clearer order so I move Year and Month column names to the first two columns. Additionally, I'd like to recreate column names as Chicken's different parts. So I use pivot_wider function to rearrange the data.

```{r}
# tidy data, use Chicken's different parts as column names
df <-poultry %>% select(Year, Month, Product, Price_Dollar) %>% pivot_wider(names_from = Product, values_from = Price_Dollar)
df
```


## Identify variables that need to be mutated

I found I may mutate the average price of every part of Chicken per year.

I grouped the data by year and mutate 5 columns to calculate the mean of every part of Chiken every year.



```{r}
# mutate data
df %>% 
  group_by(Year) %>% 
  mutate(whole_ave = mean(Whole),
         breast_ave= mean(`B/S Breast`),
         bone_in_ave = mean (`Bone-in Breast`),
         whole_legs_ave = mean (`Whole Legs`),
         thighs_ave = mean(`Thighs`)
         )%>%
  select(Year, Month, Whole, whole_ave, `B/S Breast`, breast_ave, `Bone-in Breast`, bone_in_ave, `Whole Legs`, whole_legs_ave, Thighs, thighs_ave)
```


