---
title: "Challenge1 Erika Nagai"
author: "Erika Nagai"
desription: "Reading in data and creating a post"
date: "09/13/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

# Packages 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(summarytools)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to read in and explore the data 'birds.csv'

```{r}
# 1. Read in the data

bird_data <- read_csv('_data/birds.csv')
bird_data

```

## Describe the data

It consists of 3077 rows and 14 columns.
This "bird" data documents the number of five livestock birds (chickens, ducks, geese and guinea fowls, pigeons, turkeys) of each year from 1960 to 2018 of 248 countries/regions. 

```{r}
#| label: summary

print(summarytools::dfSummary(bird_data),
      varnumbers = FALSE,
      plain.ascii  = FALSE,
      style        = "grid",
      graph.magnif = 0.80,
      valid.col    = FALSE,
      method = 'render',
      table.classes = 'table-condensed')
```


This data looks tidy, however, the column "Area" is not well organized as it has the name of the countries and the regions. For example, there are "Africa" and "Algeria" for the Area column, but "Algeria" is a part of "Africa".

```{r}
bird_data %>% 
  select(Area) %>%
  arrange(Area) %>%
  unique()
```
It seems like that the Area consists of 
- [1-220] 220 countris (Afghanistan - Zimbabwe)
- [221] World
- [222-248] 27 regions

```{r}
unique(bird_data$Area)
```


