---
title: "Challenge 2"
author: "Tracy Tien"
desription: "Data wrangling: using group() and summarise()"
date: "What is time these days?"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - faostat
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

-   FAOSTAT_livestock.csv

I will read in the data and do a series of "poking around" (commented out in the code below) to get a sense of the dataset, and return a more concise version of the data.

```{r}
FAOstat <- read_csv("_data/FAOSTAT_livestock.csv")
# I'm doing a series of "poking around" after reading in the data (commented out) to get a sense of the dataset.
# spec(FAOstat)
# dim(FAOstat) showed that FAOSTAT_livestock has 82116 records and 14 rows. 
# colnames(FAOstat)
head(FAOstat)
#install.packages("dplyr")
# I want to keep the descriptive columns instead of their corresponding codes.
codevars <- names(FAOstat) %in% c("Domain Code", "Area Code", "Element Code", "Item Code", "Year Code")
FAOStat_NoCodes <- FAOstat[!codevars]
FAOStat_NoCodes
```

## Describe the data

The FAOSTAT_livestock data contains information for 253 United Nations Food & Agriculture Organization countries, on livestock animals such as asses (donkeys?), buffaloes, camels, cattle, goats, horses, mules, pigs, and sheep (measured in heads.) The temporal coverage is from 1961-2018. The data were likely reported by the country (based on the "Flag Description" field denoting as "offical data") or estimated by the FAO ("FAO estimate".)


```{r}
#| label: summary
n_distinct(select(FAOStat_NoCodes, "Area"))
# table(select(FAOStat_NoCodes,"Element"))
table(select(FAOStat_NoCodes, "Item"))
table(select(FAOStat_NoCodes, "Year"))
```

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

I first grouped the data by Item (which denotes what type of livestock) and found the average for each, then I sorted them in descending order (using tally and sort). I was able to show for each country and each item, then it occurred to me that it would be interesting to just see the top producing livestock for each country. This then allows us to see the difference between the overall trend for all FAO countries, and how countries differ from that (even by how much.) Future questions might be to compare by year.

```{r}
# make a cleaner data to make a summary table, this can also be done with select()
codevars1 <- names(FAOstat) %in% c("Domain", "Domain Code","Area Code", "Element", "Element Code", "Item Code", "Unit", "Year Code", "Flag", "Flag Description")
FAOStat_cleaned <- FAOstat[!codevars1]
FAOStat_cleaned

# summary of cleaner data
library(summarytools)
print(dfSummary(FAOStat_cleaned))

# group by the type of livestock (= Item)
by_item <- FAOStat_cleaned %>% group_by(Item)

#by_item %>% tally(sort = TRUE) 
#by_item %>%
  #summarize(avg_value = mean(Value, na.rm=TRUE)) 


#For each country, group by item, and show the top average value livestock for each country.
item_by_area <- FAOStat_cleaned %>%
  group_by(Area, Item) %>%
  summarize(mean_value = mean(Value, na.rm =TRUE),
            median_value = median(Value, na.rm = TRUE)) %>%
  top_n(1)

item_by_area


```

### Explain and Interpret

Please see above section.