---
title: "Challenge 1 "
author: "Xiaoyan Hu"
description: "Reading in data and creating a post"
date: "09/16/2022??"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds.
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`

```{r}
setwd("/Users/cassie199/Desktop/22fall/DACSS601/601_Fall_2022/posts/_data")
data1 <- read.csv("/Users/cassie199/Desktop/22fall/DACSS601/601_Fall_2022/posts/_data/railroad_2012_clean_county.csv")
head(data1)
data2<- read.csv("/Users/cassie199/Desktop/22fall/DACSS601/601_Fall_2022/posts/_data/birds.csv")

```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

1. data railroad shows a statistic of railroad with state, counties and total employees
 the table contains 2930 obervations and 3 varibales
2.birds file described with informations of birds with their areas, value, items and so on. This file contains 30977 observations and 14 variables


```{r}
#| label: railroad

dim(data1)
colnames(data1)
head(data1)

#|birds file described with informations of birds with their areas, value, 
#|the data contains 30977 obersvations and 14 variables

head(data2)
colnames(data2)
dim(data2)


```
