---
title: "Challenge 1 Instructions"
author: "Connor Skowyra"
desription: "Reading in data and creating a post"
date: "09/22/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

## Railroad Data

> library(readr)
> railroad_2012_clean_county <- read_csv("posts/_data/railroad_2012_clean_county.csv")

## Birds Data

> library(readr)
> birds.csv <- read_csv("posts/_data/birds.csv")

##FAOSTAT Data

> library(readr)
> FAOSTAT_cattle_dairy.csv <- read_csv("posts/_data/FAOSTAT_cattle_dairy.csv")

> library(readr)
> FAOSTAT_country_groups.csv <- read_csv("posts/_data/FAOSTAT_country_groups.csv")

> library(readr)
> FAOSTAT_egg_chicken.csv <- read_csv("posts/_data/FAOSTAT_egg_chicken.csv")

> library(readr)
> FAOSTAT_livestock.csv <- read_csv("posts/_data/FAOSTAT_livestock.csv")

## Wild Birds Data

> library(readxl)
> wild_bird_data <- read_excel("posts/_data/wild_bird_data.xlsx")
> View(wild_bird_data)

## State County 2012 Data

> library(readxl)
> StateCounty2012<-read_excel("posts/_data/StateCounty2012.xls")

```{r}

```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

## Connor's Answer to Describing the Data

The data for all sets were either created in Excel or a plain text format called CSV (Comma Separated Values). To fully understand the datasets, first, we can make it easier by selecting specific columns using select().

An example is: select(railroad_2012_clean_county, contains("state"))

To make it easier to view the StateCounty2012 dataset, we can rename column ...2 to State to understand what values being looked at.

  ## Rename ...2 to State
  
State <- select(StateCounty2012,...2)

Once we rename our column, we will make a table to interpret our data.

  ## Make Table of State

table(State)

This table will summarize the amount of workers on the railroad per county in 2012.

```{r}
#| label: summary

```
