---
title: "Homework 3"
author: "Sanjana Jhaveri"
date: "12/07/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - homework_3
  - nfl_2019
  
  ---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(readr)
nfl <- read_csv("_data/nfl2019.csv")
View(nfl)

```

```{r}
nfl <- read_csv("_data/nfl2019.csv",
                          skip = 1,
                          col_names = c("delete", "Opponent", "Home_Ranking", "Opponent_Ranking", "Home_1st_Downs", "Home_Total_Yards", "Home_Passing_Yards", "Home_Rushing_Yards", "Home_Turnovers", "Opponent_1st_Downs", "Opponent_Total_Yards","Opponent_Passing_Yards", "Opponent_Rushing_Yards", "Opponent_Turnovers", "Home_Offensive_Ranking", "Home_Defensive_Ranking", "Home_Special_Teams", "Home", "Year", "Winner")) %>%
  select(!starts_with("delete")) %>%
  na_if("Skipped")
```
# Clean Data:
I needed to clean the data by renaming the columns to make more sense about which team's values I'm looking at which is why each column except for the year and the winner has wither home or opponent attached to it. It was getting a little confusing without knowing which team had which ranking or which team got a certain number of passing yards.
Clean data by assigning a number to each team name and a column with a W or L depending on which team wins and potential research questions include will the team with the most passing yards always win. Will the team with the most receiving yards always win? Will the team with a higher offensive rating. Rename columns to clean data as well. Will the team with the higher ranking let's say >10 always win?
```{r}
mean(nfl$Home_Total_Yards)
mean(nfl$Opponent_Total_Yards)
```

```{r}
mean(nfl$Home_Passing_Yards)
mean(nfl$Opponent_Passing_Yards)
```

```{r}
mean(nfl$Home_Rushing_Yards)
mean(nfl$Opponent_Rushing_Yards)
```

```{r}
nfl%>%
  ggplot(aes(x=`Home_Passing_Yards`, y=`Year`)) +
  geom_point()
```

```{r}
nfl%>%
  ggplot(aes(x=`Opponent_Passing_Yards`, y=`Year`)) +
  geom_point()
```

```{r}
nfl$home_winner <- with(nfl, ifelse(Home == Winner, 'Win', 'Loss'))

view(nfl)

```


#If Condtion I want to do:

If opp == winning team then make new column with value W
Else if home  == winning team then make new column with value L