---
title: "HW2"
output: html_document
author: "Xiaoyan Hu"
desription: "home work 2"
date: "10/10/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
---

```{r setup, include=FALSE}
#| label: setup


library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```
##  {.tabset}
### Homework 2 requirments 
For this homework, your goal is to read in a more complicated dataset. Please use the category tag "hw2" as well as a tag for the dataset you choose to use.
~   Read in a dataset from the _data folder in the course blog repository, or choose your own data. If you decide to use one of the datasets we have provided, please use a challenging dataset - check with us if you are not sure. 
~   Clean the data as needed using dplyr and related tidyverse packages.
~   Provide a narrative about the data set (look it up if you aren't sure what you have got) and the variables in your dataset, including what type of data each variable is. The goal of this step is to communicate in a visually appealing way to non-experts - not to replicate r-code.
~   Identify potential research questions that your dataset can help answer.

### Read in and data description

This data is a free data set I download online and its includes the data of child and adult health care quality measurement. 

```{r}
# read in data
getwd()
data1<- read_csv("/Users/cassie199/Desktop/22fall/DACSS601/601_Fall_2022/posts/healthcare.csv")
#overview of the data
summary(data1)



```

### tidy data

You can also embed plots, for example:

```{r pressure, echo=FALSE}
group_by(data1, "State")
%>%

```
:::
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### plot

