---
title: "HW 3"
author: "Saisrinivas Ambatipudi"
date: "2022-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod) 
library(gridExtra)
library(ggplot2) 
library(Hmisc) 
library(plotly) 
library(ggplotlyExtra)
library(tidyverse)
library(tidyquant)
library(timetk)
library(xts)
library(reshape2)
library(dplyr)
library(tidyr)
```

## Visualizations on the dataset.

First we need use the libraries to get the correct visualization


## Now we define the data and get the plots:

Extracting the dataset.
```{r}
crypto <- read_csv("D:/UMass/Sem 1/DACSS601/601_Fall_2022/posts/_data/crypto_maa2.csv")
crypto
```
## Summary of the dataset.

We need convert the dataset column formats to be appropriate for getting results.

```{r}
crypto$Date = as.Date(crypto$Date, format = "%d/%m/%Y")
crypto$Open = as.numeric(gsub(pattern = ',','',crypto$Open))
crypto$High = as.numeric(gsub(pattern = ',','',crypto$High))
crypto$Low = as.numeric(gsub(pattern = ',','',crypto$Low))
crypto$Close = as.numeric(gsub(pattern = ',','',crypto$Close))
crypto$Volume = as.numeric(gsub(pattern = ',','',crypto$Volume))
crypto$Marketcap = as.numeric(gsub(pattern = ',','',crypto$Marketcap))
crypto$Year = as.numeric(format(crypto$Date,"%Y"))
crypto$Currency = capitalize(crypto$Currency)
summary(crypto)
crypto

```

Now that we have done that we need to vizualize the cryptocurrencies.

```{r}
plot_ly(as.data.frame(crypto$Currency),labels = crypto$Currency,type = 'pie', hole = 0.3) %>%
  layout(xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```


Now we need to convert the data into time series for plotting.

```{r}
Bitcoin = subset(crypto,crypto$Currency == 'Bitcoin')
Bitcoin$Currency = NULL
colnames(Bitcoin) = c('BitcoinDate','BitcoinOpen','BitcoinHigh','BitcoinLow','BitcoinClose','BitcoinVolume','BitcoinMarketcap', 'BitcoinYear')
Bitcoin = xts(Bitcoin[,-1], order.by=as.Date(Bitcoin$BitcoinDate))

Litecoin = subset(crypto,crypto$Currency == 'Litecoin')
Litecoin$Currency = NULL
colnames(Litecoin) = c('LitecoinDate','LitecoinOpen','LitecoinHigh','LitecoinLow','LitecoinClose','LitecoinVolume','LitecoinMarketcap', 'LitecoinYear')
Litecoin = xts(Litecoin[,-1], order.by=as.Date(Litecoin$LitecoinDate))

XRP = subset(crypto,crypto$Currency == 'XRP')
XRP$Currency = NULL
colnames(XRP) = c('XRPDate','XRPOpen','XRPHigh','XRPLow','XRPClose','XRPVolume','XRPMarketcap', 'XRPYear')
XRP = xts(XRP[,-1], order.by=as.Date(XRP$XRPDate))

Dogecoin = subset(crypto,crypto$Currency == 'Dogecoin')
Dogecoin$Currency = NULL
colnames(Dogecoin) = c('DogecoinDate','DogecoinOpen','DogecoinHigh','DogecoinLow','DogecoinClose','DogecoinVolume','DogecoinMarketcap', 'DogecoinYear')
Dogecoin = xts(Dogecoin[,-1], order.by=as.Date(Dogecoin$DogecoinDate))

Monero = subset(crypto,crypto$Currency == 'Monero')
Monero$Currency = NULL
colnames(Monero) = c('MoneroDate','MoneroOpen','MoneroHigh','MoneroLow','MoneroClose','MoneroVolume','MoneroMarketcap', 'MoneroYear')
Monero = xts(Monero[,-1], order.by=as.Date(Monero$MoneroDate))

Stellar = subset(crypto,crypto$Currency == 'Stellar')
Stellar$Currency = NULL
colnames(Stellar) = c('StellarDate','StellarOpen','StellarHigh','StellarLow','StellarClose','StellarVolume','StellarMarketcap', 'StellarYear')
Stellar = xts(Stellar[,-1], order.by=as.Date(Stellar$StellarDate))

Tether = subset(crypto,crypto$Currency == 'Tether')
Tether$Currency = NULL
colnames(Tether) = c('TetherDate','TetherOpen','TetherHigh','TetherLow','TetherClose','TetherVolume','TetherMarketcap', 'TetherYear')
Tether = xts(Tether[,-1], order.by=as.Date(Tether$TetherDate))

NEM = subset(crypto,crypto$Currency == 'NEM')
NEM$Currency = NULL
colnames(NEM) = c('NEMDate','NEMOpen','NEMHigh','NEMLow','NEMClose','NEMVolume','NEMMarketcap', 'NEMYear')
NEM = xts(NEM[,-1], order.by=as.Date(NEM$NEMDate))

Ethereum = subset(crypto,crypto$Currency == 'Ethereum')
Ethereum$Currency = NULL
colnames(Ethereum) = c('EthereumDate','EthereumOpen','EthereumHigh','EthereumLow','EthereumClose','EthereumVolume','EthereumMarketcap', 'EthereumYear')
Ethereum = xts(Ethereum[,-1], order.by=as.Date(Ethereum$EthereumDate))

IOTA = subset(crypto,crypto$Currency == 'IOTA')
IOTA$Currency = NULL
colnames(IOTA) = c('IOTADate','IOTAOpen','IOTAHigh','IOTALow','IOTAClose','IOTAVolume','IOTAMarketcap', 'IOTAYear')
IOTA = xts(IOTA[,-1], order.by=as.Date(IOTA$IOTADate))

EOS = subset(crypto,crypto$Currency == 'EOS')
EOS$Currency = NULL
colnames(EOS) = c('EOSDate','EOSOpen','EOSHigh','EOSLow','EOSClose','EOSVolume','EOSMarketcap', 'EOSYear')
EOS = xts(EOS[,-1], order.by=as.Date(EOS$EOSDate))

TRON = subset(crypto,crypto$Currency == 'TRON')
TRON$Currency = NULL
colnames(TRON) = c('TRONDate','TRONOpen','TRONHigh','TRONLow','TRONClose','TRONVolume','TRONMarketcap', 'TRONYear')
TRON = xts(TRON[,-1], order.by=as.Date(TRON$TRONDate))

Chainlink = subset(crypto,crypto$Currency == 'Chainlink')
Chainlink$Currency = NULL
colnames(Chainlink) = c('ChainlinkDate','ChainlinkOpen','ChainlinkHigh','ChainlinkLow','ChainlinkClose','ChainlinkVolume','ChainlinkMarketcap', 'ChainlinkYear')
Chainlink = xts(Chainlink[,-1], order.by=as.Date(Chainlink$ChainlinkDate))

Cardano = subset(crypto,crypto$Currency == 'Cardano')
Cardano$Currency = NULL
colnames(Cardano) = c('CardanoDate','CardanoOpen','CardanoHigh','CardanoLow','CardanoClose','CardanoVolume','CardanoMarketcap', 'CardanoYear')
Cardano = xts(Cardano[,-1], order.by=as.Date(Cardano$CardanoDate))

Cosmos = subset(crypto,crypto$Currency == 'Cosmos')
Cosmos$Currency = NULL
colnames(Cosmos) = c('CosmosDate','CosmosOpen','CosmosHigh','CosmosLow','CosmosClose','CosmosVolume','CosmosMarketcap', 'CosmosYear')
Cosmos = xts(Cosmos[,-1], order.by=as.Date(Cosmos$CosmosDate))

Solana = subset(crypto,crypto$Currency == 'Solana')
Solana$Currency = NULL
colnames(Solana) = c('SolanaDate','SolanaOpen','SolanaHigh','SolanaLow','SolanaClose','SolanaVolume','SolanaMarketcap', 'SolanaYear')
Solana = xts(Solana[,-1], order.by=as.Date(Solana$SolanaDate))

Polkadot = subset(crypto,crypto$Currency == 'Polkadot')
Polkadot$Currency = NULL
colnames(Polkadot) = c('PolkadotDate','PolkadotOpen','PolkadotHigh','PolkadotLow','PolkadotClose','PolkadotVolume','PolkadotMarketcap', 'PolkadotYear')
Polkadot = xts(Polkadot[,-1], order.by=as.Date(Polkadot$PolkadotDate))

Uniswap = subset(crypto,crypto$Currency == 'Uniswap')
Uniswap$Currency = NULL
colnames(Uniswap) = c('UniswapDate','UniswapOpen','UniswapHigh','UniswapLow','UniswapClose','UniswapVolume','UniswapMarketcap', 'UniswapYear')
Uniswap = xts(Uniswap[,-1], order.by=as.Date(Uniswap$UniswapDate))

Aave = subset(crypto,crypto$Currency == 'Aave')
Aave$Currency = NULL
colnames(Aave) = c('AaveDate','AaveOpen','AaveHigh','AaveLow','AaveClose','AaveVolume','AaveMarketcap', 'AaveYear')
Aave = xts(Aave[,-1], order.by=as.Date(Aave$AaveDate))

```

And then we combine the time series data so that we can get a combined graph to visualize the opening prices of the cryptocurrencies in a graph.

```{r}
combined_df <- merge(Bitcoin, Ethereum, by = "Date")
combined_df <- merge(combined_df, Aave, by = "Date")
combined_df <- merge(combined_df, Cardano, by = "Date")
combined_df <- merge(combined_df, Chainlink, by = "Date")
combined_df <- merge(combined_df, Cosmos, by = "Date")
combined_df <- merge(combined_df, Dogecoin, by = "Date")
combined_df <- merge(combined_df, EOS, by = "Date")
combined_df <- merge(combined_df, Ethereum, by = "Date")
combined_df <- merge(combined_df, IOTA, by = "Date")
combined_df <- merge(combined_df, Litecoin, by = "Date")
combined_df <- merge(combined_df, Monero, by = "Date")
combined_df <- merge(combined_df, NEM, by = "Date")
combined_df <- merge(combined_df, XRP, by = "Date")
combined_df <- merge(combined_df, Polkadot, by = "Date")
combined_df <- merge(combined_df, Stellar, by = "Date")
combined_df <- merge(combined_df, Tether, by = "Date")
combined_df <- merge(combined_df, TRON, by = "Date")
combined_df <- merge(combined_df, Uniswap, by = "Date")
combined_df <- merge(combined_df, XRP, by = "Date")

combined_df[is.na(combined_df)] <- 0

ggplot(combined_df, aes(x = index(Bitcoin), y = BitcoinOpen/100, color = "Bitcoin")) +
  geom_line() +
  ggtitle("Crypto Opening price series") + 
  theme(legend.position = "top") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(
  x = "Date",
  y = "Opening Prices",
  color = "Currencies",
  caption = "Sai") +
  scale_x_date(date_labels = "%b %y", date_breaks = "6 months") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.key.width = unit(2,"cm"))+
  geom_line(aes(y = EthereumOpen, color = "Ethereum"))+
geom_line(aes(y = AaveOpen, color = "Aave"))+
geom_line(aes(y = CardanoOpen, color = "Cardano"))+
geom_line(aes(y = ChainlinkOpen, color = "Chainlink"))+
geom_line(aes(y = CosmosOpen, color = "Cosmos"))+
geom_line(aes(y = DogecoinOpen, color = "Dogecoin"))+
geom_line(aes(y = EOSOpen, color = "EOS"))+
geom_line(aes(y = IOTAOpen, color = "IOTA"))+
geom_line(aes(y = LitecoinOpen, color = "Litecoin"))+
geom_line(aes(y = MoneroOpen, color = "Monero"))+
geom_line(aes(y = NEMOpen, color = "NEM"))+
geom_line(aes(y = PolkadotOpen, color = "Polkadot"))+
geom_line(aes(y = StellarOpen, color = "Stellar"))+
geom_line(aes(y = TetherOpen, color = "Tether"))+
geom_line(aes(y = TRONOpen, color = "TRON"))+
geom_line(aes(y = UniswapOpen, color = "Uniswap"))+
geom_line(aes(y = XRPOpen, color = "XRP"))
```

Now we can create a graph for market capitalization.

```{r}
ggplot(combined_df, aes(x = index(Bitcoin), y = BitcoinMarketcap/100000000, color = "Bitcoin")) +
  geom_line() +
  ggtitle("Crypto Market Cap") + 
  theme(legend.position = "top") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(
  x = "Date",
  y = "Total Mean Capital",
  color = "Currencies",
  title = "Crypto Market Capitalization",
  subtitle = "in 100 millions", 
  caption = "Sai") +
  scale_x_date(date_labels = "%b %y", date_breaks = "6 months") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.key.width = unit(2,"cm"))+
  geom_line(aes(y = EthereumMarketcap/100000000, color = "Ethereum"))+
geom_line(aes(y = AaveMarketcap/100000000, color = "Aave"))+
geom_line(aes(y = CardanoMarketcap/100000000, color = "Cardano"))+
geom_line(aes(y = ChainlinkMarketcap/100000000, color = "Chainlink"))+
geom_line(aes(y = CosmosMarketcap/100000000, color = "Cosmos"))+
geom_line(aes(y = DogecoinMarketcap/100000000, color = "Dogecoin"))+
geom_line(aes(y = EOSMarketcap/100000000, color = "EOS"))+
geom_line(aes(y = IOTAMarketcap/100000000, color = "IOTA"))+
geom_line(aes(y = LitecoinMarketcap/100000000, color = "Litecoin"))+
geom_line(aes(y = MoneroMarketcap/100000000, color = "Monero"))+
geom_line(aes(y = NEMMarketcap/100000000, color = "NEM"))+
geom_line(aes(y = PolkadotMarketcap/100000000, color = "Polkadot"))+
geom_line(aes(y = StellarMarketcap/100000000, color = "Stellar"))+
geom_line(aes(y = TetherMarketcap/100000000, color = "Tether"))+
geom_line(aes(y = TRONMarketcap/100000000, color = "TRON"))+
geom_line(aes(y = UniswapMarketcap/100000000, color = "Uniswap"))+
geom_line(aes(y = XRPMarketcap/100000000, color = "XRP"))
```

These are just some of the visualizations that we need to work on find out more in-depth reasoning for a few of the research questions.

One way we can further visualize is to get the prices of the cryptocurrencies for each year the open, close, high and low value alone with the volume and the market capitalization.

