---
title: "Challenge 4"
author: "Janhvi Joshi"
desription: "More data wrangling: pivoting"
date: "11/10/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
  - abc_poll
  - eggs
  - fed_rates
  - hotel_bookings
  - debt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.xlsx or organiceggpoultry.xls⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions.xlsx ⭐⭐⭐⭐⭐

```{r}
hotels_booking <- read_csv("_data/hotel_bookings.csv")
hotels_booking
```

### Briefly describe the data
This dataset summarises various details about a hotel booking and contains 120k records and 32 columns dating from year 2015 to 2017. There are two types of hotels in this dataset - Resort Hotel and City Hotel. Customers from around the world; around 160-170 countries, book these hotels. It can also be seen that on an average, around 37% of the bookings are cancelled and around 3% of the guests are repeated. On an average, customers may need to wait for 2.3 days in the waitlist to confirm a booking and around 57% of these bookings include some special requests. The hotels provide 4 different types of meals to their customers. This dataset is likely gathered from various online and offline channels - from where the booking was made.

## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

I looked at multiple different columns and found a few that have many NULL values as shown below. So I will be removing rows where value of country is NULL since that record does not have much significance. The company columns also has many NULL values but since it is numerical, I will change it to NA. Additionally, I am not sure what the adr column depicts and will be removing it too since it is not useful for my analysis.

```{r}
table(hotels_booking$company)
table(hotels_booking$country)

```
```{r}
hotels_booking <- hotels_booking %>% 
  filter(!(country == "NULL"))
```

Any additional comments?

## Identify variables that need to be mutated

Now that null values from countries has been removed, we can see that the company column also has a lot of NULL values. In addition to that, I am going to combine the day, month and year of arrival of all guests. This will help in better analysis of the data. As mentioned above, I am also removing the adr column.

Document your work here.

```{r}
hotels_booking_mutate <- hotels_booking %>% 
  mutate(arrival_date = str_c(arrival_date_day_of_month,
                              arrival_date_month,
                              arrival_date_year, sep="/"),
         arrival_date = dmy(arrival_date)) %>% 
  select(-c(arrival_date_day_of_month,arrival_date_month,arrival_date_year, adr))

hotels_booking_mutate
```
```{r}
summary(hotels_booking_mutate$arrival_date)
```
Any additional comments?

We can see that the above mutation can help hotels in further analysing their busiest days and prepare accordingly. Next, I will change the class of the company column from character to numeric.

```{r}
hotels_booking_mutate <- hotels_booking_mutate %>%
  mutate(across(c(company),~ replace(.,str_detect(., "NULL"), NA))) %>% mutate_at(vars(company),as.numeric)
is.numeric(hotels_booking_mutate$company)
```