---
title: "Challenge 1"
author: "Karla Barrett-Dexter"
description: "Bird Data Summary"
date: "9/25/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

```{r}
birds <- read_csv("_data/birds.csv")
birds
```



## Describe the data
It appears the data is showing the price per 1000 head of different types of birds.

Basic information about the data is as follows:
There are 5 different categories of birds represented in the data set. 
There are 248 countries represented in the data set.
The data was collected over 58 years, 1961-2018.



```{r}
#| label: summary

birds %>% count(Item) #Used to find the summarize the categories of birds and total birds per category
Countries<- birds%>%count(Area) #Practiced creating a variable
count(Countries) #Used to find the total number of countries represented in the data
Years <- birds %>% count(Year)#I used this code to create a cleaner list of the years represented in the data
first(Years) #I used this code to order the years from earliest to latest


```



Using filters to drill down Armenia's data: 
There are 2 different categories of birds, Chickens and Turkeys.
Armenia data spans 27 years, 1992-2018.

```{r}
#| label: Armenia
ArmeniaBirds <- filter(birds, Area == "Armenia")
ArmeniaBirds
ArmeniaBirds %>% count(Item)
ArmeniaYears <- ArmeniaBirds %>% count(Year)
first(ArmeniaYears)


```
