---
title: "Challenge 1 Instructions"
author: "Meredith Rolfe"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
library(readxl)
railroad<-read_csv("_data/railroad_2012_clean_county.csv")
railroad
```

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}

#| label: summary
#| 

```

# Railroad Dataset(railroad_2012_clean_county.csv)-

There are total three variables- state, county, and total employees. We will write R commands to see how many states, county and total employees unique employees are there and we will observe the dataset accordingly.

```{r}
railroad%>%
  select(county)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(county)%>%
  distinct()
```

```{r}
railroad%>%
  select(state)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(state)%>%
  distinct()
```

```{r}
railroad%>%
  select(total_employees)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(total_employees)%>%
  distinct()
```

# Birds Dataset(FAOstat\*.csv)-

There are total three variables- state, county, and total employees. We will write R commands to see how many states, county and total employees unique employees are there and we will observe the birds dataset accordingly.

```{r}
library(readxl)
birds<-read_csv("_data/birds.csv")
birds
```

```{r}
chickens<-read_csv("_data/FAOSTAT_egg_chicken.csv")
chickens
```

```{r}
birds.sm<-birds%>%
  select(-contains("Code"))
birds.sm
```

# Wild Birds Dataset(wild_bird_data.csv)-

```{r}
wildbirds<-read_excel("_data/wild_bird_data.xlsx")
wildbirds
```

```{r}
wildbirds <- read_excel("_data/wild_bird_data.xlsx", skip = 1)
wildbirds
```

Here we can see wet body weight and population size, now we will clean up the dataset little bit.

```{r}
wildbirds <- read_excel("_data/wild_bird_data.xlsx", skip = 2, col_names = c("weight", "pop_size"))
wildbirds
```

# Railroad Dataset(wild_bird_data.csv)-

There are total three variables- state, county, and employees. We will write R commands to see how many states, county and employees unique employees are there and we will observe the wild_bird_data dataset accordingly.

```{r}
read_excel("_data/StateCounty2012.xls",skip = 3)
```

Let's delete the unwanted columns and make it clean and select the on which we need in the use. Then we will filter out the total rows and remove any notes if necessary.

```{r}
read_excel("_data/StateCounty2012.xls",skip = 4,col_names= c("State", "delete", "County", "delete", "Employees"))%>%
  select(!contains("delete"))
```

```{r}
railroad<-read_excel("_data/StateCounty2012.xls", skip = 4,col_names= c("State", "delete", "County", "delete", "Employees"))%>%
  select(!contains("delete"))%>%
  filter(!str_detect(State, "Total"))
railroad
```

```{r}
railroad <-head(railroad, -2)
tail(railroad, 10)
```

```{r}
railroad%>%
  select(State)%>%
  distinct()
```

```{r}
railroad%>%
  select(State)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(County)%>%
  distinct()
```

```{r}
railroad%>%
  select(County)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(Employees)%>%
  distinct()
```

```{r}
railroad%>%
  select(Employees)%>%
  n_distinct(.)
```
