---
title: "Challenge 1 Instructions"
author: "Meredith Rolfe"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories: 
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
library(readxl)
railroad<-read_csv("_data/railroad_2012_clean_county.csv")
railroad
```

# Railroad Dataset-

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
railroad%>%
  select(county)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(county)%>%
  distinct()
```

```{r}
railroad%>%
  select(state)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(state)%>%
  distinct()
```

```{r}
railroad%>%
  select(total_employees)%>%
  n_distinct(.)
```

```{r}
railroad%>%
  select(total_employees)%>%
  distinct()
```

# Birds Dataset-

```{r}
library(readxl)
birds<-read_csv("_data/birds.csv")
birds
```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.
