---
title: "Homework 2"
author: "Matthew O'Neill"
desription: "MA School Distict Data Exploration"
date: "12/01/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - Matthew O'Neill
  - MA_Public_Schools_2017
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

For this Homework I decided to use the dataset I plan to work with for the final project. It contains extensive information about all schools in the state from the year 2017. For this assignment, I plan to explore the dataset, build a narrative around it, and begin to identify research questions which could be solved, at least in part, with this data.

## Reading in Data


```{r}

data<- read_csv("_data/MA_Public_Schools_2017.csv",show_col_types = FALSE)

head(data, 10)


```


The first step is to read in the data. I printed out the first 10 lines of data to look through and see what we have to work with.

The dataset has 302 columns, which is great. We might not need all of them though, specifically the MCAS scores for different grades accross different skill areas. These might be useful to us in future assignments or in the final project, but for now, I think I'm more interested in High School statistics.

## Narrative

This is a comprehensive dataset of every state funded public school in Massachusetts for the fiscal year 2017. Each row of data represents a school and each school has a large variety of useful data. Below is a breakdown of that data.


  * Financial Data
    + Teacher Salaries
    + Total Expenditures
    + Expenditures Per Pupil
  * Socioeconomic Makeup of Student Body
    + Percentage of students of different racial background
    + Percentage of economically disadvantaged students
    + Percentage of Disabled studnets
    + Percentage Male vs Female Students
  * Enrollment Data
    + Enrollment by Grade
    + Average Class Size
  * Test Scores
    + SAT Scores w/ Subject Breakdown
    + MCAS Scores w/ Subject Breakdown
    + AP Scores and Number of AP Test Takers
  * High School Outcome
    + Graduation Rate
    + Dropout Rate
    + Rate of Secondary Education
    
It's important to discuss some limitations that this dataset has. 

Although it is very compehensive, it is relatively old. However, the questions I envision myself attemtping to research are not very time sensitive and could still produce useful results.

The data is also specific to Massachusetts Public Schools, which means there will be relatively fewer schools with extreme class sizes, like you might find in an area with very high population density. There will also be relatively higher average salaries and expenditures accross the board, since Massachusetts is one of the wealthiest states in the country.

That said, value can still be extracted from researching this data as it can inform local education decision making.



## Potential Research Questions

I chose this dataset in particular because at one time, I planned to become a high school Math teacher. I was very interested in the impact that salary and class size had on the academic outcomes of students, and how much access students from different socioeconomic backgrounds had to higher quality education.

Two formal research questions I plan to answer in the final project are: 

* How does average class size and average distict expenditure impact graduation rates and rates of enrollment in higher education?
* Do studnets in different socioeconomic backgrounds have differing levels of access to smaller class sizes and high expenditure?
    




