---
title: "Homework 2"
author: "Michaela Bowen"
desription: ""
date: "10/12/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - Michaela Bowen
  - dataset
  - ggplot2
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(readr)
library(summarytools)
library(lubridate)

knitr::opts_chunk$set(echo = TRUE)
```


:::panel-tabset
### Read In Data
This is the current cannabis inventory. This is including products that have been sold out or contain 0 quantities in order to have the THC and expiration date information for products that have already been sold out.

Below I have read in the current cannabis inventory. There are 6741 instances of 12 variables, meaning these current products are on the Sales Floor Room, Quarantine room, or vault. I have renamed "ID" columns `sku`, and `product ID` to "delete" and selected relevant columns. I have separated the THC column into the numeric value and a column that contains the unit of that value. For some non-cannabis items, "expiration date" is unnecessary and those values of "invalid date" have been changed to `NA`. I then arranged the data by room, category and product. The last column, `isCannabis` has been categorized and ordered as a boolean because it is a yes or no variable. The `price`, `cost`, and `thc` columns are numeric. The expiration date column was separated and formated according to the date and date. All other columns are character columns
```{r, message = FALSE, warning= FALSE}
Current_Inventory <- read_csv("../posts/2022-11-16 Export inventory.csv",
                                         skip = 1,
                                         col_names = c("delete","product","package ID","batch","available","room","price","cost","category","expiration_date","thc","isCannabis"))%>%
  select(!contains("delete"))%>%
  separate(col = "thc", into = c("thc","unit_thc (%/mg)"), sep = " ")%>%
  mutate(thc = as.numeric(thc),
        isCannabis = case_when(
           isCannabis == "Yes" ~ TRUE,
           isCannabis == "No" ~ FALSE
         ))%>%
  mutate(expiration_date = na_if(expiration_date, "Invalid date"),
         expiration_date = mdy(expiration_date))%>%
  mutate(
    date = as.Date(expiration_date),
    day = day(expiration_date))%>%
  mutate(
    format_date = format(date, "%m/%d/%Y")
  )%>%
  arrange(desc(room), category, product)

```
:::








