---
title: "Challenge 2 Instructions"
author: "Meredith Rolfe"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - faostat
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readxl)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xls ⭐
-   FAOstat\*.csv or birds.csv ⭐⭐⭐
-   hotel_bookings.csv ⭐⭐⭐⭐

Here I have read in the RailRoads Excel File Sheet. Extraneous and empty data has been removed, and Canada had its county named "Canada". 


```{r}
railroads <-read_excel("_data/StateCounty2012.xls",
                     skip = 4,
                     n_max = 2987,
                     col_names= c("state", "delete",  "county",
                                  "delete", "employees"))%>%
  select(!contains("delete"))%>%
  filter(!str_detect(state, "Total"))%>%
  mutate(county = ifelse(state=="CANADA", "CANADA", county))

railroads

```

## Describe the data

The variables are named *State*, *County*, and *Employees*. This is likely the data of railroad employees working at counties within the United States. There are 54 distinct "States" which include the District of Columbia and Canada, the other two are likely oversea or military. 

```{r}
#| label: summary
railroads%>%
  select(state)%>%
  n_distinct(.)


railroads%>%
  select(state)%>%
  unique()

```

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

```{r}

```

### Explain and Interpret

Be sure to explain why you choose a specific group. Comment on the interpretation of any interesting differences between groups that you uncover. This section can be integrated with the exploratory data analysis, just be sure it is included.
