---
title: "Challenge 1 Instructions"
author: "Neeharika Karanam"
desription: "Reading in data and creating a post"
date: "09/24/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

```{r}
library(readxl)
wild_birds_dataset<-read_excel("_data/wild_bird_data.xlsx", skip=1)
wild_birds_dataset
```
## Display the dimensions of the data

```{r}
dim(wild_birds_dataset)
```

The data present in the form of an excel has been read and stored into the variable(wild_birds_dataset). The first line of the dataset consists of the description of where it has been taken from which can be ignored for our analysis. Therefore, I have skipped the first row of the dataset and then read the excel and displayed the details. The dataset consists of 2 columns the "body weight" and the "population_size". The dimensions of the dataset is displayed i.e [146, 2].

## Describe the data

```{r}
#Checking if there are any null values
is.null(wild_birds_dataset)
```

```{r}
#Summary of the dataset
summary(wild_birds_dataset)
```

The wild birds dataset consists of only 2 columns body_weight and population_size which are both numerical values. Before we actually perform any analysis on the data and try to understand the distribution of the values in the dataset we need to make sure that there are no NULL values present in the data as it might effect our data adversely. We Could clearly see that there are no NULL values in the dataset we can summarize the data which will provide us  the statistical data of the dataset. The birds which have a very low body weight are very less in number whereas the birds which are heavier and have high body weight are high in number. This clearly says that the surival rate or the population size is directly proportional to the body weigt of the bird. The mean, median, min, max and other statistical information is displayed above.