---
title: "Challenge 2 Instructions"
author: "Nikita Masanagi"
desription: "Data wrangling: using group() and summarise()"
date: "11/05/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - faostat
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xls ⭐
-   FAOstat\*.csv or birds.csv ⭐⭐⭐
-   hotel_bookings.csv ⭐⭐⭐⭐

```{r}
# Import important libraries needed for functions
library(tidyverse)
library(rmarkdown)
library(summarytools)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r}
# Read data from the csv file
data <- read.csv('_data/hotel_bookings.csv')

# Printing the read data as a paged table
paged_table(data)

```
This dataset shows the records of hotel bookings across different countries with details like booking data, booking status, no of adults/children/babies, etc.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
# Get and print summary of the dataset to help us analyze it.
print(dfSummary(data))

```

As can be seen in the above summary, hotel has only two values - Resort and City. The dataset has data from the year 2015 to the year 2017. It can be observed that about 37% of the bookings are canceled from the is_canceled column summary. We can see a large number of columns with categorical values like country, market_segment, distribution_channel, customer_type, etc. and how the percentage of these categorical values are distributed in the data set. This gives us insight on what analysis can be done on the data.

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.


```{r}
# Group dataset by Market Segment
dataGroupedByMS <- data %>%
  group_by(market_segment)

# Analyze the grouped segment's stays in weekdays and weekends using mean of the number of nights.
dataGroupedByMS %>%
  summarise(
    avg_stays_in_week_nights = mean(stays_in_week_nights, nr.rm = TRUE),
    avg_stays_in_weekend_nights = mean(stays_in_weekend_nights, nr.rm = TRUE)
  )

```

### Explain and Interpret

The first table gives insight into which market segment bookings come for weekdays and which ones come during weekends. This could help the hotels decide which events to organize or how to plan activities targeted to a certain segment during those days.

The second table gives percentage of bookings of each type of segment that were cancelled, no-shows or checked-out. This could help the hotels analyze which market segments gets the most cancellations and which ones get the least, and find ways to reduce the number of cancellations by targeted schemes or penalties towards certain segments.


