---
title: "Challenge 7 "
author: "Nikita Masanagi"
description: "Visualizing Multiple Dimensions"
date: "08/24/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_7
  - hotel_bookings
  - australian_marriage
  - air_bnb
  - eggs
  - abc_poll
  - faostat
  
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  Recreate at least two graphs from previous exercises, but introduce at least one additional dimension that you omitted before using ggplot functionality (color, shape, line, facet, etc) The goal is not to create unneeded [chart ink (Tufte)](https://www.edwardtufte.com/tufte/), but to concisely capture variation in additional dimensions that were collapsed in your earlier 2 or 3 dimensional graphs.
   - Explain why you choose the specific graph type
5) If you haven't tried in previous weeks, work this week to make your graphs "publication" ready with titles, captions, and pretty axis labels and other viewer-friendly features

[R Graph Gallery](https://r-graph-gallery.com/) is a good starting point for thinking about what information is conveyed in standard graph types, and includes example R code. And anyone not familiar with Edward Tufte should check out his [fantastic books](https://www.edwardtufte.com/tufte/books_vdqi) and [courses on data visualizaton.](https://www.edwardtufte.com/tufte/courses)

(be sure to only include the category tags for the data you use!)

## Read in data

Choosing the debt data from last week

```{r}
RawData <- read_excel("_data/debt_in_trillions.xlsx")
head(RawData)
```

### Briefly describe the data

The data represents the year and quarter in the US and the associated debt spread of its citizens during that time period.

## Tidy Data (as needed)

Pivoting the data here to filter by debt type

```{r}
longerSplitData<- splitData%>%
  pivot_longer(!c(Year,Quarter), names_to = "DebtType",values_to = "DebtPercent" )

longerSplitData

```

## Visualization with Multiple Dimensions

Original second graph from last challenge

```{r}
longerSplitDataPlot <- longerSplitData%>%
  ggplot(mapping=aes(x = Year, y = DebtPercent))


longerSplitDataPlot + 
  geom_point(aes(color = DebtType))
```
Only the value of the total and mortgage is shown,rest of the data is lost.Hence we can improve the chart to show debt types.

```{r}
longerSplitDataPlot + 
  geom_point(aes(color = DebtType))+
  labs(title = "Total National Debt most closely follows Mortgages",subtitle="Student and Auto drive the secondary category" ,caption = "data from posts/_data")+
  theme(legend.position = "bottom")+
  facet_zoom(y = DebtType == !c("Mortgage","Total"),ylim = c(0,2))
```
Original third graph from previous challenge

It is difficult to observe the patterns without changing the scales.We can adjust the axis to see how thr debt flows through the years.
```{r}
longerSplitDataPlot+
  geom_point() +
  facet_wrap(~DebtType) +
  scale_x_discrete(breaks = c('03','06','09',12,15,18,21))

```
```{r}
longerSplitDataPlot+
  geom_point(aes(color = Quarter,alpha=0.9,)) +
  facet_wrap(~DebtType,scales = "free_y") +
  scale_x_discrete(breaks = c('03','06','09',12,15,18,21))+
  theme_light() +
  guides(alpha="none") +
  labs(title = "Debt Types per year" ,caption = "data from posts/_data")
```




