---
title: "Challenge 8"
author: "Nikita Masanagi"
description: "Joining Data"
date: "08/25/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_8
  - railroads
  - snl
  - faostat
 
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in multiple data sets, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  join two or more data sets and analyze some aspect of the joined data

(be sure to only include the category tags for the data you use!)

## Read in data

Using the FAOSTAT dataset

```{r}
codes <- read_csv("_data/FAOSTAT_country_groups.csv")
cattle <- read_csv("_data/FAOSTAT_cattle_dairy.csv")
```



### Briefly describe the data

The primary dataset I have chosen is the Fao Stat Cattle dataset.It has information about available food and agriculture data from 245 countries.
It contains information about cow milk, units sold and price of the product.the data ranges from 1960,s to 2018.

The next file I will using to perform the join is codebook, which groups the data by countries to provide a high level overview, instead of at the country level.

We can then perform join within these datasets and analyze within the country groups.


## Tidy Data (as needed)

Matching the column name of the two tables, as one has Area code and one has Country Code. Renaming the cattle Area Code with Country Code.

```{r}
cattlenew <- rename (cattle, "Country Code"= "Area Code" )
head(cattlenew)


```



## Join Data

We can perform left join by using the Country Code

```{r}
cattlefinal <- left_join(cattlenew, codes, by = "Country Code" )
head(cattlefinal)
```
After performing join, we can now display the graph by Country Code.

```{r}
cattlefinal %>%
  filter(Year >= 2000) %>%
  filter(`Flag Description` == "Official data") %>%
  filter(`Country Group`=="Central Asia") %>%
  filter(`Unit` == "tonnes") %>%
  ggplot(aes(x=`Area`, y=`Value`)) + 
  geom_point(
    color="blue",
    fill="#69b3a2",
    size=.5,
    alpha=.5
  )+
  labs(title = "Official Tonnes of Milk Produced in Central Asia from the Year 2000-2018", x="Country", y="Price of Milk") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) 
```



