---
title: "Challenge 1 Instructions"
author: "Nikita Masanagi "
desription: "Reading in data and creating a post"
date: "09/24/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}

#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
library(readxl)
wildbird <- read_excel("_data/wild_bird_data.xlsx",skip=1)
wildbird

```

```{r}
dim(wildbird)

```
Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: summary

summary(wildbird)
```

```{r}
is.null(wildbird)
```

The data has been extracted paper by Nee et al and there are 146 rows and 2 columns. The first column indicates wet body difference and the second column indicates the population of the birds in the wild. Both columns have floating point values. The data set does not have missing values or null values but when running the summary function we get two NA’s as the as.numeric function converted the first row in numeric values. The minimum value is 5 and the maximum value is 5093378 The mean and median is 191619 and 491 respectively.
