---
title: "Challenge 8 Instructions"
author: "Prajakti Kapade"
description: "Joining Data"
date: "08/25/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_8
  - railroads
  - snl
  - faostat
  - debt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in multiple data sets, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  join two or more data sets and analyze some aspect of the joined data

(be sure to only include the category tags for the data you use!)

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

  - military marriages ⭐⭐
  - faostat ⭐⭐
  - railroads  ⭐⭐⭐
  - fed_rate ⭐⭐⭐
  - debt ⭐⭐⭐
  - us_hh ⭐⭐⭐⭐
  - snl ⭐⭐⭐⭐⭐

```{r}
data_1 <- read_csv('_data/FAOSTAT_livestock.csv')
data_2 <- read_csv('_data/FedFundsRate.csv')
data_1
data_2
```

### Briefly describe the data
The dataset1 shows the information about Livestock statistics with the Area, Domain and Items according to the year.
The dataset2 shows the information about the fed funds rate,showing inflation rate, GDP, unemployement etc for each year.
## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

The data is tidy already, so we can work directly with it.

Are there any variables that require mutation to be usable in your analysis stream? For example, do you need to calculate new values in order to graph them? Can string values be represented numerically? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

Document your work here.

```{r}
joined_data <- merge(data_1,data_2,on='year')
joined_data
```

## Join Data
The joined data shows the livestock and fed rates, and shows how the rates have affected the livestock in that year.

Be sure to include a sanity check, and double-check that case count is correct!

```{r}
vis1 <- joined_data %>% 
  group_by(Year, Item) %>%
  summarize(
    n = n()
  )
ggplot(vis1, aes(fill=Item, y=n, x=Year)) + 
  geom_bar(stat="identity")+ #sta="identity" is necessary
  labs(title = "Change in the number of livestocks over the Years")
```
I have analyzed the change in number of livestocks based on the Items in the data along the years.
```{r}
ggplot(data = joined_data, aes(x = Unemployment.Rate)) +
      theme(axis.text.x = element_text(angle = 90))+
  geom_bar(stat = "count",fill='lightpink') +
  labs(title = "Records by Unemployment rate ",
       x = "Number of Records",
       y = "Unemployment rate")
```
I have also analyzed the trend of Unemployment rate.
