---
title: "HW2"
author: "Prajakti Kapade"
description: "Homework 2"
date: "10/26/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true

---

```{r}
library(dplyr) 
```
The columns of the dataset are :
```{r}
data_country <- read_csv('_data/FAOSTAT_livestock.csv')
colnames(data_country)
data_country
```
The columns/features are of the following type:
1. Domain.Code - categorical variable, related to the Domain
2. Domain - categorical variable
3. Area.Code - Categorical/Ordinal, related to the Area
4. Area - categorical variable
5. Element.Code - Categorical variable related to the element
6. Element - Categorical variable
7. Item.Code - Categorical variable related to the Item
8. Item - Categorical variable
9. Year.Code  - Categorical variable related to the Year (it is similar to the Year)
10. Year - Categorical variable
11. Unit - categorical
12. Value - Continuous variable
13. Flag - categorical related to Flag.Description
14. Flag Description - Categorical

As, we can see some columns are related to each other we can choose certain columns from it, to reduce complexity.
We are taking a small subset of the data by extracting some columns of use.
```{r}
data_country_subset <- data_country[c('Domain','Area','Element','Item','Year','Unit','Value','Flag.Description','Flag')]
data_country_subset
```

The summary of the columns in the dataset are shown below:
```{r}

#install.packages( pkgs = "summarytools" )
summary(data_country)
```

From the code below, we can see only one element is present.
```{r}
date_country_groups<-group_by(data_country,Element)
abc_data.d_by_product<-summarise(date_country_groups,
          n=n())
abc_data.d_by_product
```

```{r}
date_country_groups<-group_by(data_country,Item)
abc_data.d_by_product<-summarise(date_country_groups,
          n=n())
abc_data.d_by_product
```

There is only one domain present in the dataset.
```{r}
date_countryg<-group_by(data_country,Domain)
abc_data.d_by_product<-summarise(date_countryg,
          n=n())
abc_data.d_by_product
```
Research Question: 
1. Which area has higher number of livestocks, what are the median, mean and mode values for the same? 
2. Across which years were the count of livestocks the highest? 
3. What is the unit type across areas and which one is the highest in numbers?
4. How is flag affecting the livestock value count?

