---
title: "Challenge 4 Instructions"
author: "Prajakti Kapade"
desription: "More data wrangling: pivoting"
date: "08/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
  - abc_poll
  - eggs
  - fed_rates
  - hotel_bookings
  - debt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.xlsx or organiceggpoultry.xls⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions.xlsx ⭐⭐⭐⭐⭐

```{r}
abc_data <- read.csv('_data/poultry_tidy.csv')
abc_data


```

### Briefly describe the data
The data consists of the Product and the price in dollar. It is based on the data collected from Poultry farms. It is intended to collect data about each type of Poultry and their time period along with the Price in dollars.
## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

I formatted the data to combine the year and month of the data using the following command, to get better grouping based on months and years together.

```{r}
abc_data$Date <- format(as.Date(paste0(abc_data$Month, abc_data$Year, "01"), format="%b%Y%d"), "%m-%Y")
abc_data

```

Any additional comments?

## Identify variables that need to be mutated

Are there any variables that require mutation to be usable in your analysis stream? For example, are all time variables correctly coded as dates? Are all string variables reduced and cleaned to sensible categories? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

I have grouped the data based on Product and the Date field, which we have build using Year and Month. It gives the mean and median Price in dollars for each category. As, the points are very less due to lesser categories, it could not be visualized well, but for larger datasets this would work very well.

Document your work here.

```{r}
product_groups<-group_by(abc_data,Product)
abc_data.p_by_product<-summarise(product_groups,
          p_mean=mean(Price_Dollar),
          p_median=median(Price_Dollar),
          n=n())
abc_data.p_by_product<-arrange(abc_data.p_by_product,Product)
head(abc_data.p_by_product)
```

```{r}
ggplot(aes(Product,p_mean),data=abc_data.p_by_product) +
 geom_point()
```
```{r}
date_groups<-group_by(abc_data,Date)
abc_data.d_by_product<-summarise(date_groups,
          p_mean=mean(Price_Dollar),
          p_median=median(Price_Dollar),
          n=n())
abc_data.d_by_product<-arrange(abc_data.d_by_product,Date)
head(abc_data.d_by_product)
```
```{r}
ggplot(aes(Date,p_mean),data=abc_data.d_by_product) +
 geom_point()
```
Any additional comments?
