---
title: "Challenge 7 Instructions"
author: "Prajakti Kapade"
description: "Visualizing Multiple Dimensions"
date: "08/24/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_7
  - hotel_bookings
  - australian_marriage
  - air_bnb
  - eggs
  - abc_poll
  - faostat
  - usa_households
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  Recreate at least two graphs from previous exercises, but introduce at least one additional dimension that you omitted before using ggplot functionality (color, shape, line, facet, etc) The goal is not to create unneeded [chart ink (Tufte)](https://www.edwardtufte.com/tufte/), but to concisely capture variation in additional dimensions that were collapsed in your earlier 2 or 3 dimensional graphs.
   - Explain why you choose the specific graph type
5) If you haven't tried in previous weeks, work this week to make your graphs "publication" ready with titles, captions, and pretty axis labels and other viewer-friendly features

[R Graph Gallery](https://r-graph-gallery.com/) is a good starting point for thinking about what information is conveyed in standard graph types, and includes example R code. And anyone not familiar with Edward Tufte should check out his [fantastic books](https://www.edwardtufte.com/tufte/books_vdqi) and [courses on data visualizaton.](https://www.edwardtufte.com/tufte/courses)

(be sure to only include the category tags for the data you use!)

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

  - eggs ⭐
  - abc_poll ⭐⭐
  - australian_marriage ⭐⭐
  - hotel_bookings ⭐⭐⭐
  - air_bnb  ⭐⭐⭐
  - us_hh ⭐⭐⭐⭐
  - faostat ⭐⭐⭐⭐⭐

```{r}
data <- read_csv('_data/hotel_bookings.csv')
data
```

### Briefly describe the data
This dataset contains the data for different types of hotels along with the information about the stays, the duration and the number of people staying (adults and childern). This gives an analysis of the pattern in hotel industry.
## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

```{r}
data$dateval <- paste(data$arrival_date_year,data$arrival_date_month)
data$dateval
```

Are there any variables that require mutation to be usable in your analysis stream? For example, do you need to calculate new values in order to graph them? Can string values be represented numerically? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

Document your work here.

```{r,fig.width=11,fig.height=5}
library(ggplot2)
library(dplyr)

# Most basic bubble plot
ggplot(data = data, aes(x = paste(arrival_date_year,arrival_date_month), y = stays_in_weekend_nights)) +
      geom_point(aes(color = as.factor(market_segment))) +
      labs(title = "Stay in weekend nights based on month and year ",
           x = "Month/Year", y = "Stays in weekend",angle = 45,width = 1000, height=1000)
```

## Visualization with Multiple Dimensions
```{r,fig.width=8,fig.height=4}


# Most basic bubble plot
ggplot(data = data, aes(x = hotel, y = adults, fill = as.factor(market_segment))) +
      geom_bar(stat='identity') +
      labs(title = "Number of adults for hotel type for each market segment ",
           x = "Hotel Types", y = "Number of adults",angle = 45)
```


