---
title: "Challenge 2"
author: "Saisrinivas Ambatipudi"
desription: "Data wrangling: using group() and summarise()"
date: "10/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroads
  - faostat
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xls ⭐
-   FAOstat\*.csv or birds.csv ⭐⭐⭐
-   hotel_bookings.csv ⭐⭐⭐⭐

```{r}
hotel1 <- read_csv("_data/hotel_bookings.csv")
hotel1
```

Add any comments or documentation as needed. More challenging data may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

After using the read_csv function we can see that the data has 119320 rows and 32 columns. After using the summary function we can see that:
- The dates span from 2015 to 2017 for this data set.
- In one booking 55 adults booked a room when we check the max of the adults column.
- In one booking 10 children and in another booking 10 babies based on max values of the children and the babies column respectively.
- A booking was on waiting list for a maximum of 391 days!!
- The maximum number of 5 special requests was made on one booking.
- The maximum number of 8 parking spaces were requested on one booking, looks like it was the time 55 people came to the place.XD

```{r}
summary(hotel1)
```

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

```{r}
hotel2 <- hotel1 %>% group_by(hotel, arrival_date_month, arrival_date_year)
hotel3 <- hotel2 %>% summarise(Babies = sum(babies), Children = sum(children, na.rm = TRUE), Adults = sum(adults))
hotel3



x <- filter(hotel3, arrival_date_year == 2015)
y <- filter(hotel3, arrival_date_year == 2016)
z <- filter(hotel3, arrival_date_year == 2017)


x1 <- filter(x, hotel == "City Hotel")
x1
y1 <- filter(y, hotel == "City Hotel")
y1
z1 <- filter(z, hotel == "City Hotel")
z1

x2 <- filter(x, hotel == "Resort Hotel")
x2
y2 <- filter(y, hotel == "Resort Hotel")
y2
z2 <- filter(z, hotel == "Resort Hotel")
z2


summary(x)
summary(y)
summary(z)
summary(x1)
summary(y1)
summary(z1)
summary(x2)
summary(y2)
summary(z2)

sum(hotel3$Babies, na.rm=TRUE)
sum(hotel3$Children, na.rm=TRUE)
sum(hotel3$Adults, na.rm=TRUE)
sum(x1$Babies, na.rm = TRUE)
sum(x1$Children, na.rm = TRUE)
sum(x1$Adults, na.rm = TRUE)
sum(y1$Babies, na.rm = TRUE)
sum(y1$Children, na.rm = TRUE)
sum(y1$Adults, na.rm = TRUE)
sum(z1$Babies, na.rm = TRUE)
sum(z1$Children, na.rm = TRUE)
sum(z1$Adults, na.rm = TRUE)
sum(x2$Babies, na.rm = TRUE)
sum(x2$Children, na.rm = TRUE)
sum(x2$Adults, na.rm = TRUE)
sum(y2$Babies, na.rm = TRUE)
sum(y2$Children, na.rm = TRUE)
sum(y2$Adults, na.rm = TRUE)
sum(z2$Babies, na.rm = TRUE)
sum(z2$Children, na.rm = TRUE)
sum(z2$Adults, na.rm = TRUE)


```

### Explain and Interpret

Be sure to explain why you choose a specific group. Comment on the interpretation of any interesting differences between groups that you uncover. This section can be integrated with the exploratory data analysis, just be sure it is included.

So the data analysis that I have done is in regard to how many babies, children and adults stayed in each hotel for the three different years which are city hotel and resort hotel for the years 2015, 2016 and 2017 respectively.

From the analysis we can see that for city hotel the analysis is:
          
Babies: 2015 ==> 84, 2016 ==> 195, 2017 ==> 113    
Children: 2015 ==> 512, 2016 ==> 3814, 2017 ==> 2922    
Adults: 2015 ==> 24385, 2016 ==> 70568, 2017 ==> 51885                                                 


For the resort hotel it is:

Babies: 2015 ==> 139, 2016 ==> 845, 2017 ==> 15877   
Children: 2015 ==> 268, 2016 ==> 2094, 2017 ==> 34396    
Adults: 2015 ==> 150, 2016 ==> 2216, 2017 ==> 24525 


When we take the analysis for both hotels together, the total number of babies, children and adults are 949, 12403 and 221636 respectively.

The minimum and maximum number of babies for the city hotel in a month for the years are:

City Hotel:                    

Max   2015 ==>23, 2016 ==>27, 2017 ==>19
Min   2015 ==> 2, 2016 ==> 7, 2017 ==>10                 

Resort Hotel:
Max   2015 ==>51, 2016 ==>55, 2017 ==>30
Min   2015 ==>11, 2016 ==> 7, 2017 ==>7                

The minimum and maximum number of children for the city hotel in a month for the years are:

City Hotel: 

Max   2015 ==> 165, 2016 ==> 730, 2017 ==> 604
Min   2015 ==> 14, 2016 ==>  69, 2017 ==> 165                  

Resort Hotel:
Max   2015 ==> 263, 2016 ==> 438, 2017 ==> 580
Min   2015 ==>  47, 2016 ==> 47, 2017 ==> 92                

The minimum and maximum number of adults for the city hotel in a month for the years are:

City Hotel:                     
Max   2015 ==> 6282, 2016 ==> 7755, 2017 ==> 8287
Min   2015 ==> 1918, 2016 ==> 2200, 2017 ==> 4178                

Resort Hotel:
Max   2015 ==> 3230, 2016 ==> 3664, 2017 ==> 3659
Min   2015 ==> 1796, 2016 ==> 1496, 2017 ==> 2150                 

        







