---
title: "HW2_Saisrinivas"
author: "Saisrinivas Ambatipudi"
date: "2022-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Reading in Data:

The data we will read in will be a specific cryptocurrency data and we will describe it.

```{r cars}
x <- read_csv("D:/UMass/Sem 1/DACSS601/601_Fall_2022/posts/_data/crypto_maa2.csv")
x

```

## Summary of the data

From the above data we will now take the summary function and describe it.

```{r pressure, echo=FALSE}
summary(x)
```

We can see that the there 8 columns in the dataset.
Currency, date, open, high, low, close, volume and marketcap.
Currency is the name of the cryptocurrency, it class is character.
Date is the column which give us the date of the cryptocurrency so its a recurring value. Its class is also character but we will need to convert it into date class because it ill help us create a time series to plot the graph later.
Open is the column which tells us the opening price on that day. It is a numeric class.
High is the highest value of the currency on that day. It is a numeric class.
Low is the lowest value of that currency on that day. It is a numeric class.
Close is the value of the currency after the market closes on that day. It is a numeric class.
Volume is the total amount of coins that are available for trade. It is a numeric class.
Market cap is the value of the coins multiplied by the volume. It is a numeric class.

## Potential Research Questions:

Some of the research questions we can use to analyze or find something from this data is:

- Which coin dominated in which years in terms of market capitalization?
- Why the different prices of the cryptocurrencies from the years 2017 to 2021 because thats when these digital assets became a super hit?
- Why Bitcoin's volume decreased during 2020-2021 while tether's volume kept increasing?


