---
title: "Sarah McAlpine - Challenge 2"
author: "Sarah McAlpine"
desription: "Reading in Data and Summarizing Its Groups"
date: "09/20/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - sarahmcalpine
  - cattle
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(tidyr)
library(summarytools)
library(readr)

knitr::opts_chunk$set(echo = TRUE)
```

## Challenge Overview

Today's challenge is to read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc.) Next, provide summary statistics for different interesting groups within the data, and interpret those statistics.

## Reading in the Data
I will read in the FAO cattle data for my first attempt and may return to this to do another.


```{r}
#read in the data. Use read_csv over read.csv
cattle<- read_csv("_data/FAOSTAT_cattle_dairy.csv")
colnames(cattle)

#preview data without the 100% values
head(select(cattle, Area, Element, Year, Unit, Value))%>%
  group_by(Area) %>%
slice(1:10)

# sort out the 3 different Elements into their own columns?

```
## Refining Distinct Values
Next I will print a summary with only the columns that have multiple distinct values.
```{r}
cattle.sm <- select(cattle, Area, Element, Year, Unit, Value)

# preview with data frame summary
print (dfSummary (cattle.sm, varnumbers = FALSE,
                plain.ascii = FALSE,
                style = "grid" ),
      method = 'render', 
      table.classes = 'table-condensed')
```

## Rearranging the Data
I see that there are 3 "Elements": Milk Animals, Production, and Yield; and 3 "Units" and I suspect these are aligned where each element is measured in its own unit. In order to get a better view of this data and ensure that each column contains only one factor, I will move those into their own columns with `pivot_wider'. See my attempt below.
```{r}
# Move the units into their own columns, pulling their values along. 

#use read_csv over read.csv
pivot_wider(cattle.sm, id_cols = c(Area, Year), 
            names_from = Unit, 
            values_from = Value)%>%
  # Rename columns to describe the moved variables
  rename(Head_Cattle = Head, Animal_hg = "hg/An", Tonnes_Milk = tonnes)%>%
 #Stop showing the elements, now that they are tied to the units in the new column names
   select(Area, Year, Head_Cattle, Animal_hg, Tonnes_Milk) %>%
  #attempt to remove the NAs and collapse the rows together. Didn't work--need to try again next time.
  head()
 #would I use case_when for collapsing the rows together? 
#How do I assign a new name to this format of the data? Where can I put that command, near the beginning?

```
## Reading and Tidying Together
Now that I have explored the data, I'm ready to set up my read-in with all the tidying all at once so that I can work with a single data set for my analysis. Below is my script to do so. Since all the steps are piped in one script, the assignment of "clean.dairy" applies to the new, tidy data set. 
```{r}
library(tidyverse)
clean.dairy <- read_csv("_data/FAOSTAT_cattle_dairy.csv") %>%
  filter(Flag!="A")%>%

  select(Area, Year, Unit, Value) %>%

  pivot_wider(names_from = Unit, 
            values_from = Value) %>%
  # Rename columns to describe the moved variables
  rename(Head_Cattle = Head, hg_Animal = "hg/An", Tonnes_Milk = tonnes)
#preview to confirm
head(clean.dairy)
```
## Summary Statistics
As shown below, average (mean) values of these three variables ??what do I observe??

```{r}

  
  #the following works for a table of min, max, avg for Head_Cattle, hg_Animal, and Tonnes_Milk.
clean.dairy %>%
  select(clean.dairy, Area = China)
   summarize(avg_Head_Cattle = mean (Head_Cattle, na.rm=TRUE),
          med_Head_Cattle = median (Head_Cattle, na.rm=TRUE), 
          max_Head_Cattle = max (Head_Cattle, na.rm=TRUE),
          min_Head_Cattle = min(Head_Cattle, na.rm=TRUE))
          
 clean.dairy %>%
  summarize(avg_hg_Animal = mean (hg_Animal, na.rm=TRUE),
          med_hg_Animal = median (hg_Animal, na.rm=TRUE), 
          max_hg_Animal = max (hg_Animal, na.rm=TRUE),
          min_hg_Animal = min(hg_Animal, na.rm=TRUE))   
 
 clean.dairy %>%

  summarize(avg_Tonnes_Milk = mean (Tonnes_Milk, na.rm=TRUE),
          med_Tonnes_Milk = median (Tonnes_Milk, na.rm=TRUE), 
          max_Tonnes_Milk = max (Tonnes_Milk, na.rm=TRUE),
          min_Tonnes_Milk = min(Tonnes_Milk, na.rm=TRUE))  

 #Counts of unique occurences
 distinct(clean.dairy, Year)

  distinct(clean.dairy, Area)

```
## Summary
The FAOSTAT data on cattle contains 58 years of data from 1961 to 2018 on cattle meat and milk production for 204 countries (232 minus 28 "aggregate" regions). 

For my analysis, I am interested in the top producers of meat and milk. I have relocated the "Items" of production to their own columns for a tidy set of data. From there, I can observe the max, min, and averages of heads of cattle, meat production, and milk production. The top 10 producers of Meat in hectograms is shown below, and of milk in tonnes below that. Interestingly, Canada is the only country that appears in both lists. 

As a side note, I am not sure why Northern America appears in the top beef producers table, since I removed aggregate values when I read in the data. Perhaps there was an error there.

In the future, it may be interesting to compare rates of meat and milk production in proportion to the head of cattle. Another question to investigate could be comparing countries' production vs consumption of cattle products.

```{r}
# Top 10 Beef Producers in a single year
clean.dairy %>%
select(Area, hg_Animal)%>%
   group_by(Area) %>%
  arrange(desc(hg_Animal)) %>%
            slice(1)%>%
  ungroup()%>%
  arrange(desc(hg_Animal)) %>%
  slice(1:10)

# Top 10 Milk Producers in a single year
clean.dairy %>%
select(Area, Tonnes_Milk)%>%
   group_by(Area) %>%
  arrange(desc(Tonnes_Milk)) %>%
            slice(1)%>%
  ungroup()%>%
  arrange(desc(Tonnes_Milk)) %>%
  slice(1:10)
```

