---
title: "Sarah McAlpine HW 2"
author: "Sarah McAlpine"
desription: "Homework 2"
date: "10/11/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - sarahmcalpine
  - inaturalist data
  - lubridate
  - summarytools
  - time zones
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(lubridate)
library(summarytools)

knitr::opts_chunk$set(echo = TRUE)
```

## Select a Data Set

For this assignment, I chose to use a set of data from iNaturalist.org of citizen scientist observations of plant life using mobile apps. This particular data set is limited to those observations made in North America whose identifications have the most disagreements. The iNaturalist site allows for custom data queries, but since I am not familiar with each data field, I exported more than I needed. First I will read in this data and take a look at the column names and values present within each to decide what to keep. At the beginning, I have 27 columns and 25,266 rows.

At the outset, possible research questions could be about patterns in family, order, genus, and/or species that are difficult for the general public to identify, and possibly certain areas of the continent that have the highest likelihood of identification disagreement. I would not, however, be able to compare which users are most likely to have disagreed idetntifications since my data doesn't include undisputed identifications as well.

```{r}
#read in data
observations <- read_csv("_data/plant_observations.csv")

# preview data and plan for cleaning
 colnames(observations)
 head(observations)
 print(dfSummary(observations,
     varnumbers = FALSE,
     plain.ascii  = FALSE, 
     style = "grid",
     graph.magnif = 0.60, 
     valid.col    = FALSE),
      method = 'render',
      table.classes = 'table-condensed')


#[1] "id"           DELETE              "observed_on"  1973 - today        "time_observed_at"       LUBRIDATE     
#[4] "time_zone"     LUBRIDATE          "user_id"      DELETE              "user_login"             15.3k    
#[7] "created_at"     LUBRIDATE         "updated_at"   LUBRIDATE           "quality_grade"          3 categories
#10] "num_identification_agreements"    "num_identification_disagreements" "captive_cultivated"     T/F      
#13] "place_guess"       MAP?           "latitude"               MAP?      "longitude"              MAP? 
#16] "place_town_name"    MAP?          "place_state_name"      MAP?       "place_country_name"     MAP?     
#19] "species_guess"                    "scientific_name"                  "common_name"                     
#22] "iconic_taxon_name"                "taxon_id"              DELETE     "taxon_order_name"       GROUP    
#25] "taxon_family_name"     GROUP      "taxon_genus_name"      GROUP      "taxon_species_name"  

#clean up time zones

```

## Building a Clean Read-in

```{r}
#cleaning during read-in
obs_clean <- read_csv("_data/plant_observations.csv",
                      skip = 1,
                      col_names = c("id", 
                                   "delete",
                                   "time_observed_at",
                                   "delete",
                                   "delete",
                                   "user_login",
                                   "delete",
                                   "delete",
                                   "quality_grade",
                                   "num_identification_agreements",
                                   "num_identification_disagreements",
                                   "captive_cultivated",
                                   "place_guess", 
                                   "latitude",
                                   "longitude",
                                   "place_town_name",
                                   "place_state_name",
                                   "place_country_name", 
                                   "species_guess",
                                   "scientific_name",
                                   "common_name",
                                   "iconic_taxon_name",
                                   "delete",
                                   "taxon_order_name",
                                   "taxon_family_name",
                                   "taxon_genus_name",
                                   "taxon_species_name")) %>%
#exclude columns called delete
select(!starts_with("delete"))


```

## Dealing with Time Zones

After many hours, I was able to use `dplyr` to remove the time zone characters from the datetime strings, and then I was able to use `ymd_hms` to convert those to datetimes. Since I am not researching the times of day, I forced the UTC time zone for all these data. 

One other idea: could I use latitudes to calculate time zones?

\-[lubridate source sheet](https://lubridate.tidyverse.org/articles/lubridate.html)

\-[lubridate cheat sheet](https://rawgit.com/rstudio/cheatsheets/main/lubridate.pdf)

\-[tz database time zones](https://en.wikipedia.org/wiki/List_of_tz_database_time_zones)

```{r}

#strip time zone info
obs_clean$time_observed_at <- str_remove_all (obs_clean$time_observed_at, "\\W0000")

  
# turn into a date  
obs_clean$time_observed_at <- ymd_hms(obs_clean$time_observed_at,
          tz = "UTC")                                                             

#preview
head(obs_clean)
```
## Identify Research Questions
* What are some of the most disputed families, genuses of plants?
* Do these vary by location or year?

```{r}
# find top 10 families
ranked_families <-  obs_clean %>%
  select(taxon_family_name) %>%
  count(taxon_family_name) %>%
  arrange(desc(n)) %>%
mutate(prop_families = round(n/sum(n),3))


# find proportions of top 10 families
obs_clean %>%
  select(taxon_family_name) %>%
  count(taxon_family_name) %>%
  arrange(desc(n)) %>%
  slice(1:11) %>%
  mutate(prop_families = round(n/sum(n),3))
#top 10 included NA, so I changed to 11
```

