---
title: "Challenge 1 Instructions"
author: "Tejaswini_Ketineni"
desription: "Reading in data and creating a post"
date: "08/21/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

I am going to work with one data set: 
1. railroad_2012_clean_county.csv


## Reading the Data

Initially the data set-1(railroad_2012_clean_country) is read

```{r}
library(readxl)
railroad_2012_clean_county <- read_csv("_data/railroad_2012_clean_county.csv")
View(railroad_2012_clean_county)
```

Head function is used to understand the population of data 

```{r}
head(railroad_2012_clean_county)
```

```{r}
rows_and_columns_ds1 <- dim(railroad_2012_clean_county)
rows_and_columns_ds1
```
There are about 2930 rows and about 3 columns present in the dataset.

```{r}
names_col <- colnames(railroad_2012_clean_county)
names_col
```
There are three columns present in the data set namely : state,county and total_employees present


```{r}
sum(is.na(railroad_2012_clean_county))
sum(is.null(railroad_2012_clean_county))
```
There are no nulls or missing values present in the data set

```{r}
summary(railroad_2012_clean_county)
```

```{r}
library(data.table)
data_railroad <- data.table(railroad_2012_clean_county)
data_railroad[, .(distinct_states = length(unique(state)))]
data_railroad[, .(distinct_county = length(unique(county)))]
```

There are 53 distinct states and 1709 distinct counties present.

```{r}
(table(railroad_2012_clean_county$state))
```

## Description of the data

 The data set taken is analysed and the following observations are made.There are about 2930 rows and about 3 columns namely (state, county and the total_employees) present in the data set.The data set is checked for null and missing values.We observe that there are no such values present and the data set is clean.The summary statistics are checked.the count of unique states present in the data set is 53 and 1709 unique counties are present.Tabulating the states and the total_employees, we see that the highest number of employees are present in Texas(TX) and Georgia(GA) while the lowest employee count is observed in Armed forces(AE),Armed forces Pacific(AP).We also observe that the no.of states with employee count <10 is very less.

