---
title: "Challenge 4 Solutions"
author: "Tejaswini Ketineni"
date: "11/07/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
---
##Challenge_4


```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
library(stringr)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


## Read in data

```{r}
library(readr)
FedFundsRate <- read_csv("_data/FedFundsRate.csv")

```

### Briefly describe the data

To see sample columns in the data set we do a head

```{r}
head(FedFundsRate)
```
```{r}
colnames(FedFundsRate)
```

```{r}
dim(FedFundsRate)

```

```{r}
print(summarytools::dfSummary(FedFundsRate,
                        varnumbers = FALSE,
                        plain.ascii  = FALSE, 
                        style        = "grid", 
                        graph.magnif = 0.70, 
                        valid.col    = FALSE),
      method = 'render',
      table.classes = 'table-condensed')

```

## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

we see that there are many na's in the data, so we replace all of them by 0's

```{r}
FedFundsRate<- FedFundsRate %>% replace_na(list(`Federal Funds Target Rate`=0,`Federal Funds Lower Target`=0,`Inflation Rate`=0, `Federal Funds Upper Target`=0,`Real GDP (Percent Change)` =0, `Unemployment Rate` =0,`Effective Federal Funds Rate`=0))
```



## Identify variables that need to be mutated

I have observed that year month and day are separated, So it's optimal to have the date formatted.


```{r}
FedFundsRate$Date <- str_c(FedFundsRate$Year,"-",FedFundsRate$Month,"-",FedFundsRate$Day)%>%ymd()%>%as.Date()
```

```{r}
FedFundsRate$Date
```

```{r}
feds <- FedFundsRate %>% 
  pivot_longer(cols = c( "Federal Funds Target Rate","Inflation Rate","Federal Funds Lower Target",
                         "Federal Funds Upper Target","Effective Federal Funds Rate","Unemployment Rate","Real GDP (Percent Change)"),
               names_to = "names",
               values_to = "value",
               values_drop_na = TRUE)

feds

```
deleting the first three columns
```{r}
feds<-subset(feds,select=-c(1,2,3))
```

```{r}
head(feds)
```

The data is now tidy enough to work