---
title: "Challenge 1 Solution"
author: "Gabriel Katz"
description: "Reading in data and creating a post"
date: "09/21/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
  - faostat
  - wildbirds
---

```{r}

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xls ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
read_csv("railroad_2012_clean_county.csv")
```

The tibble above was read into RStudio using the "read_csv" command

While the following tibble does not focus on analyzing the whole dataframe, I have a reason for presenting this specific set of data: As a born and raised Maineiac I wanted to see how the counties stacked up in terms of the amount of railroad employees present in each one. While Aroostook county may be the largest by land area in the state, it certainly does not have the highest overall population and thus not so unexpectedly it comes in 6th in terms of the amount of railroad employee population. I was able to present this specific set of data by using the "filter" and "arrange" functions in RStudio.

```{r}
read_csv("railroad_2012_clean_county.csv") %>%
  filter(state == "ME") %>%
  arrange(desc(total_employees))

```
